### 2010-02-12 peace/papers/NAR_web_server_2009/paper.tex
Modifications to manuscript: Conclusion added; corrections made

\documentclass[a4,center,fleqn]{NAR}
\usepackage[pdftex,
colorlinks=true, %makes the links show up in color, rather than in a box
citecolor=black, %color of the in-text citation numbers
urlcolor=blue %color of the url links
]{hyperref}

\bibliographystyle{unsrt}

% Enter dates of publication
\copyrightyear{}
\pubdate{}
\pubyear{}
\jvolume{}
\jissue{}

%\articlesubtype{This is the article type (optional)}

\begin{document}

\title{PEACE: {\underline P}arallel {\underline E}nvironment for {\underline A}ssembly
  and {\underline C}lustering of Gene {\underline E}xpression}

\author{D.M. Rao\,$^{1}$, J.C. Moler\,$^{1}$, M. Ozden\,$^1$, Y. Zhang\,$^{1}$,
  C. Liang$^{1,2*}$\, and J.E. Karro\,$^{1,3}$\footnote{to whom
    correspondence should be addressed}}

\address{$^1$ Department of Computer Science and Software Engineering, \\
  $^2$ Department of Botany, \\
  $^3$ and Department of Microbiology, Miami University, Oxford, Ohio,
  USA}

\history{Received Feb. 12, 2010}

\maketitle

\begin{abstract}
  We present {\sc Peace}, a standalone tool for the high-throughput
  {\it ab initio} clustering of transcript fragment sequences produced by Next
  Generation or Sanger Sequencing technologies.  Installed and run
  through an easily downloaded GUI, {\sc Peace} can process large data
  sets and group the fragments by gene association, achieving a
  greater sensitivity (without sacrificing runtime) than the competing
  clustering tool {\sc WCD}.  Through the GUI the user can collect
  cluster statistics and examine specific clusters for more
  comprehensive study.  By using {\sc Peace} as a ``pre-assembly''
  tool to allow the application of an assembly tool to the resultant
  individual clusters, the user gains both a significant increase in
  sensitivity and a large decrease in overall runtime for the
  clustering and assembly together.
\end{abstract}

\section{Introduction}

Understanding an organism's transcriptome, the set of (spliced)
transcripts expressed by genes of the organism, is a vital step in
understanding the full functional and organizational role of the
genome in the life cycle of any organism.  Studying the transcriptome
has led to gene discovery, provided information on splice variants,
and helped shed light on the biological processes both controlling and
dictated by the genome \cite{Nagaraj07}.  But to access to those
transcripts, we must deal with the fragmented data produced by
both Next Generation and traditional Sanger sequencing technology.

Until recently, access to a transcriptome sequence was primarily
through the use of Expressed Sequence Tags (ESTs), single-pass cDNA
sequences derived from transcribed mRNAs and sequenced by Sanger
Sequencing technology.  More recently, Next Generation Sequencing
(NGS) technologies have replaced the Sanger Sequencing techniques,
allowing for considerably coverage of the transcriptome.  For example,
ESTs in GenBank dbEST are increasingly the product of NGS technologies
such as 454 pyrosequencing, which enables the sequencing of novel and
rare transcripts at a considerably higher rate of coverage than did
Sanger Sequencing \cite{Cheung2006,Emrich2007}.  From a computational
perspective, this is a mixed blessing: while NGS provides immense
quantities of new information, it also provides immensely larger data
sets, requiring fast computational techniques if we are to process
them.

Given a set of transcript fragments sampled from across the genome, a
necessary first step in the analysis is that of clustering: separating
the fragments according to the transcript from which they were
derived.  Frequently performed implicitly by assembly tools,
clustering the data as a ``pre-assembly'' step has a number of
advantages.  Most significant among these: performing this step will
allow the application of the assembly tool to individual clusters --
saving significant amounts of time due to the reduced input size
\cite{Hazelhurst08a}.  But clustering is a computationally challenging
problem.  Even with the smaller number of ESTs produced using Sanger
sequencing, the runtime and memory requirements to cluster based on
pair-wise sequence alignments make such an approach infeasible.  The
much larger data-set size produced by Next-Generation sequencing
technologies exacerbates this problem.

Here we present {\sc Peace}, a tool for the {\it ab initio} clustering
of transcript fragments by gene, applicable to both short-read and
traditional Sanger Sequencing technologies.  Available through the
\href{http://www.peace-tools.org}{www.peace-tools.org} website, the
PEACE GUI allows the user to both easily install (locally or remotely)
and run the computational tool, as well as enabling transparent
parallel processing and providing various tools for result analysis.
Using a {\it minimum spanning tree} (MST) based clustering strategy
\cite{Jain99,Wan08} and the $d^2$ sequence distance function
\cite{Hide94}, {\sc Peace} generates the MSTs using a tailored version
of Prim's algorithm \cite{Prim57}.  The result is an easy-to-use and
clustering tool that , without any sacrifice in runtime, is
considerably more sensitive and robust to sequencing errors than other
clusters tools in the literature \cite{Burke99,Slater00,
  Huang99,Parkinson02,Kalyanaraman03,Malde03,Ptitsyn05,Hazelhurst08a, Picardi09}.

\enlargethispage{-65.1pt}

\section{PEACE: Installation and Use}

The {\sc Peace} GUI, available from the {\sc Peace} website
(\href{http://www.peace-tools.org}{www.peace-tools.org}), handles the
details of both installation and job management and provides tools for
the initial analysis of the results.  By downloading and executing the
{\tt peace.jar} GUI Java program on a Linux, Windows or OS-X machine
running Java 1.6+, the user is presented with the {\sc Peace}
management tool.  Through this the user can install the computational
tool and perform a clustering of a data file, view an initial analysis
of the clusters and produce files containing subsets of the clusters
for assembly by a tool such as {\sc Cap3} \cite{Huang99}.  A typical
(first) use of the tool would generally require the the following
steps.

\noindent {\bf Tool Installation:} Installing the GUI requires only
that the Java file be downloaded from the website.  To install the
{\sc Peace} tool onto a local or remote machine, the user will select
from within the GUI the appropriate menu tab (Figure~1(a)), which
then starts the install wizard that will prompt for the appropriate
information.  Figure~1(b) illustrates the request for server
information; the user has chosen to install the {\sc Peace}
computational tool on a remote machine and is providing the necessary
connection information.  Server information is persistent between GUI
sessions, giving the user access to {\sc Peace} on the target machine
as needed.

\noindent {\bf Job Processing:} After importing the target file into
the GUI, the user can start a new job and follow the wizard menus.
Figure~1(c) illustrates the process of specifying the number of
processors available (if running on a machine supporting the OpenMPI
protocol -- which will be determined during job installation).  Once
executed, the GUI will manage the job thread, alert the user when the
job is completed (or when the user next runs the GUI after completion)
and copy the final results back to the local machine if necessary.

\noindent {\bf Result Analysis:} Once the resulting clusters have been
computed, the user has several options
for analysis:
\begin{itemize}
\item {\bf Export:} The user can export the contents of one or more
  clusters into a FASTA format file, obtaining a subset of the original file
  containing the sequences corresponding to the selected clusters
  ready for processing by an assembly tool (e.g. {\sc Cap3} \cite{Huang99}).
\item {\bf View Clustering}: The user may view a list of clusters,
  expanding selected clusters to a list of all fragments (illustrated
  in Figure~1(d)).
\item {\bf Classified Summary Graph:} The user may view a distribution
  of cluster sizes.  Further, the user may set up a {\it classifier},
  associating certain patterns with specific colors.  These patterns
  will be matched against the fragment header information from the
  original FASTA file, allowing the overlay of a colored cluster size
  distributions.  For example, if the sequence names contain unique
  string patterns denoting different cDNA libraries, the classifier
  can help the user to determine the differential expression profiles
  of different libraries for a given cluster by color coding the
  strains.  The method of setting up a these classifiers, and the
  resulting histogram, is illustrated in Figure~1(e).
\end{itemize}
Extensive documentation for the tool has been posted on the {\sc
  Peace} website, as well as links to several videos demonstrating
{\sc Peace} use and capabilities.

\section{Methods}

The clustering performed by {\sc Peace} is based on the use of minimum
spanning trees (MSTs), known to be an effective approach for narrow
band single linkage clustering \cite{Jain99,Wan08}. We use the $d^2$
distance measure \cite{Hide94} to represent edge weights.  Prim's
algorithm \cite{Prim57} then allows us to efficiently calculate an
MST from which we can infer a high-quality clustering solution.

The $d^2$ distance measure used to assign edge weights is an
alignment-free measurement of sequence distance that can be calculated
significantly faster than a Smith-Waterman alignment \cite{Hide94}.
$d^2$ works by comparing the frequency of words (strings of a fixed
length) appearing in a limited region of each string.  Fragments
overlapping by a sufficient length will share neighborhoods of enough
similarity to ensure a small distance even in the presence of a
moderate number of base errors.

We can model the fragment input as a weighted, undirected graph: the fragments
are represented as nodes, with $d^2$ sequence distances assigned to
the connecting edges as weights.  Conceptually, we want to remove each
edge exceeding a threshold score from the complete graph, and define
our partitions by the remaining connected components of the graph.  An
edge with a large weight connects fragments which are likely unrelated;
once such edges are removed, the components define a series of
overlaps.  Hence those fragments that can still be connected by some path
correspond to the same gene.  However, such an approach requires the
calculation of all edge weights.  Such a task is infeasible both in terms of
runtime and memory usage for the data set sizes we expect to process.

We approach the problem by generating a minimum spanning tree of the
described graph, then removing edges exceeding our threshold.  By
using Prim's algorithm we are able to calculate edge weights
on-the-fly (reducing memory requirements), and we can skip the
calculation of a majority of edge distances using the $u/v$ and $t/v$
filtering heuristics employed in {\sc wcd} \cite{Hazelhurst08a}.
These heuristics allow us to quickly dismiss many of the edges as too
large without the need to apply the full $d^2$ algorithm (see
Sections~A and B of the Supplementary Materials for more details).

\section{Results}

We have tested {\sc Peace} on both simulated and real data,
comparing results against those produced by the {\sc wcd}
clustering tool \cite{Hazelhurst08a} and the {\sc Cap3} assembly
tool \cite{Huang99} (which calculates a cluster as an intermediate
step).  For our simulation tests we used the {\bf ESTSim} tool
\cite{Hazelhurst03} to generate simulated transcript of varying length
fragments under different models
of error (Supplementary Materials, Section~C.1),
generating the fragments from the list of 100 zebra fish genes used in
the {\sc wcd} testing \cite{Hazelhurst08a}.  Tool parameters were
taken to match, as closely as possible, those used in the {\sc wcd}
study (see supplementary materials).  Our primary methods of quality
assessment were {\it sensitivity} (the fraction of fragment pairs from the
same gene that were correctly clustered together) and {\it Type 1
  error} (the fraction of genes that were divided between clusters)
\cite{Wang04,Hazelhurst08a}; other measurements are also discussed
in the supplementary materials.  In Figure~2(left) we see a
significant improvement of {\sc Peace} (blue) over {\sc wcd}
(green) and {\sc Cap3} (black) in its sensitivity to sequencing
errors, while in Figure~2(right) we see a comparable
improvement in Type 1 error.  In Figure~3 we look at
the number of singleton clusters (fragments not joined to any cluster),
which will not occur in our simulated sets; we again see significant
improvement in {\sc Peace}.  The sequential runtime of
{\sc Peace} is slightly improved over that of {\sc wcd} (see
Supplementary Materials, Figure~S4).

\begin{figure*}[b]
\centerline{
\includegraphics[trim=0cm 3cm 0cm 1cm, clip, scale=0.9]{screen.d/screen.pdf} 
}
\label{screen}
\caption{Screenshots of the PEACE GUI during execution, including (a)
GUI Welcome and server installation menu; (b) setup wizard for
installing the computational tool on a remote server; (c) execution
wizard for starting a selected job to be executed in parallel mode;
(d) basic cluster output; and (e) histogram view of cluster results
and classifier editor for setting up differential expression profiles.}
\end{figure*}

\begin{figure*}[b]
\centerline{
\includegraphics[scale=0.5]{pics.d/SeT1.pdf} 
}
\label{SeT1}
\caption{Comparisons of Sensitivity and Type 1 error, based on the
  average over 30 simulated fragment sets derived from 100 zebra fish genes
  (see Supplementary Materials, Section~\href{sim_results}, for more
  details).   Blue/Solid = {\sc Peace}, Green/Dash =
  {\sc wcd}, Black/Dot-Dash = {\sc Cap3}; vertical tics = 95\% confidence
  intervals on estimates.  Intervals are not presented for Type 1
  error due to the extremely small variance.}
\end{figure*}

\begin{figure*}[b]
\centerline{
\includegraphics[scale=0.35]{pics.d/singletons.pdf}
}
\label{singletons}
\caption{Average number of fragments flagged as singletons by each tool
  when run on the simulated sequences; correct answer in all cases is
  zero.  Blue/Solid = {\sc Peace}, Green/Dash = {\sc wcd},
  Black/Dot-Dash = {\sc Cap3}.}
\end{figure*}

In applying the tools to real data, we started with a set of
approximately $190,000$ clean Sanger ESTs derived from the {\it
  Chlamydomonas reinhardtii} genome \cite{Liang2008}.  To compute
sensitivity we used the {\it gmap} tool \cite{Wu05} to map the
fragment set to the genome, taking this as our reference clustering.
We see slight improvements in {\sc Peace} over {\sc wcd} in both
sensitivity and Type 1 error (both significantly outperforming {\sc
  Cap3}).  Using the mouse fragment set used in Hazelhurst {\it et
  al.}  \cite{Hazelhurst08a} {\sc Peace} again shows slight
improvements in sensitivity and Type 1 error rates, with an 18\% speedup
for {\sc Peace} (see Supplementary Materials,
Section~\href{real_results}, for more details).

\section{Conclusions}

Here we have presented {\sc Peace}, a stand-alone tool for the
high-throughput clustering of transcript fragments, managed through a
user friendly GUI that enables both local and remote instillation and
execution in sequential or parallel mode.  Based on a novel algorithm
for the clustering of the fragments by gene association, {\sc Peace}
shows significant improvement in sensitivity while increasing runtime
over the competing {\sc WCD} tool \cite{Hazelhurst08a}, and shows an
order of magnitude in improvement over the clustering preformed in the
course of assembly by the {\sc Cap3} tool \cite{Huang99}.

As a clustering tool based on sequence distance, {\sc Peace} faces
certain inherent limitations. For example, {\sc Peace} cannot handle
duplicate genes; like {\sc WCD}, it is unable to separate clusters
corresponding to genes with a greater than 88\% similarity.
Similarly, naturally biological effects such as the trans-splicing of
transcripts, or effects from poorly cleaned transcript data (e.g. the
failure to remove polyA tailed introduced in the closing process) can
cause similar effects in these clustering tools.  The problems can be
handled through the application of the assembler, and the ability to
apply any assembler to small cluster (as opposed to the data set as a
while) results in a significant reduction in overall assembly time.

In short: {\sc Peace} can out-preform existing fragment cluster tools
on fragment data.  By using the {\sc Peace} GUI, the user can easily
set up a first analysis of transcript distributions and set the user
up for a significant reduction in the overall time for assembly a
transcript from fragment transcript sequences.

\section{Acknowledgements}

Dr. Karro was funded under a PhRMA Foundation Informatics Research
Starters Grant while conducting this research.  We would also like to
acknowledge Iddo Friedberg, David Woods, Jens Mueller and David
Scoville at Miami University for their help with this project.

\vspace{3mm}
%Bib TeX

\bibliography{peace.bib}

\end{document}

% LocalWords:  arallel nalysis lustering ngine Rao Moler Ozden Zhang Liang ESTs
% LocalWords:  Karro XXXXX transcriptome pre isoforms SNPs unsequenced runtime
% LocalWords:  chlamydomondomonas ESTate PaCE xsact CLU WCD Hazelhurst et al th
% LocalWords:  CHUN nalyzer Prim's MSTs subgraph MPI SourceCode ESTSim runtimes
% LocalWords:  ests Chlamydomonas Reihartdii gmap CDNA transcriptomes cDNA www
% LocalWords:  reinhartdii conifergdb DNAc bp wcd reinhardtii PhRMA Iddo fasta
% LocalWords:  Friedberg Scoville OpenMPI initio multi Screenshots nvironment
% LocalWords:  ssembly xpression GenBank dbEST pyrosequencing ESTsim sim mRNAs
% LocalWords:  NGS polyA