### 2010-02-13 peace/papers/NAR_web_server_2009/paper.tex
Further esits to manuscript

\documentclass[a4,center,fleqn]{NAR}
\usepackage[pdftex,
colorlinks=true, %makes the links show up in color, rather than in a box
citecolor=black, %color of the in-text citation numbers
urlcolor=blue %color of the url links
]{hyperref}

\bibliographystyle{unsrt}

% Enter dates of publication
\copyrightyear{}
\pubdate{}
\pubyear{}
\jvolume{}
\jissue{}

%\articlesubtype{This is the article type (optional)}

\begin{document}

\title{PEACE: {\underline P}arallel {\underline E}nvironment for {\underline A}ssembly
  and {\underline C}lustering of Gene {\underline E}xpression}

\author{D.M. Rao\,$^{1}$, J.C. Moler\,$^{1}$, M. Ozden\,$^1$, Y. Zhang\,$^{1}$,
  C. Liang$^{1,2*}$\, and J.E. Karro\,$^{1,3}$\footnote{to whom
    correspondence should be addressed}}

\address{$^1$ Department of Computer Science and Software Engineering, \\
  $^2$ Department of Botany, \\
  $^3$ and Department of Microbiology, Miami University, Oxford, Ohio,
  USA}

\history{Received Feb. 12, 2010}

\maketitle

\begin{abstract}
  We present {\sc Peace}, a stand alone tool for the high-throughput
  {\it ab initio} clustering of transcript fragment sequences produced
  by Next Generation or Sanger Sequencing technologies.  Installed and
  run through an easily downloaded GUI, {\sc Peace} can process large
  data sets and group the fragments by gene association, achieving a
  greater sensitivity (without sacrificing runtime) than the competing
  clustering tools.  Through the GUI the user can collect
  cluster statistics and examine specific clusters for more
  comprehensive study.  {\sc Peace}, used as a ``pre-assembly'' tool,
  enables any assembly tool to work with individual small clusters of
  closely related sequences, resulting in significant increases in
  sensitive and robustness to sequencing error, as well as a reduction in
  overall runtime.
\end{abstract}

\section{Introduction}

Understanding an organism's transcriptome, the set of (spliced)
transcripts expressed by genes of the organism, is a vital step in
understanding the full functional and organizational role of the
genome in the life cycle of any eukaryote.  Studying the transcriptome
has led to gene discovery, provided information on splice variants,  
and helped shed light on the biological processes both controlling and
dictated by the genome \cite{Nagaraj07}.  However, to access those
transcripts, we must deal with the fragmented data produced by
both Next Generation and traditional Sanger sequencing technology.

Until recently, access to a transcriptome sequence was primarily
through the use of Expressed Sequence Tags (ESTs), single-pass cDNA
sequences derived from transcribed mRNAs and sequenced by Sanger
Sequencing technology.  More recently, Next Generation Sequencing
(NGS) technology has begun to rapidly replace Sanger Sequencing,
allowing for more comprehensive coverage of the transcriptome.  For
example, ESTs now being added to the GenBank dbEST are increasingly
the product of NGS technologies such as 454 pyrosequencing, which
enables the sequencing of novel and rare transcripts at a considerably
higher rate of coverage than Sanger Sequencing
\cite{Cheung2006,Emrich2007}.  From a computational perspective, this
is a mixed blessing: while NGS provides immense quantities of new
information, it also provides immensely larger data sets -- and thus a
need for faster, more efficient analysis algorithms.

Given a set of transcript fragments sampled from across the genome, a
necessary first step of their analysis is that of clustering: separating
the fragments according to the transcript from which they were
derived.  Frequently performed implicitly by assembly tools,
clustering the data as a ``pre-assembly'' step has a number of
advantages.  Most significantly: performing this step will
allow the application of the assembly tool to individual clusters --
saving significant amounts of time due to the reduced input size
\cite{Hazelhurst08a}.

But clustering is a computationally challenging problem.  Even with
the smaller number of ESTs produced using Sanger Sequencing, the
runtime and memory requirements to cluster based on pair-wise sequence
alignments make such an approach infeasible in practice.  The much
larger data-set size produced by NGS technologies exacerbates this
problem.  To deal with this, {\sc Peace} combines the $d^2$
alignment-free sequence distance function \cite{Hide94} and the
concept of a {\it minimum spanning tree} \cite{Prim57} to quickly and
accurately find clusters of ESTs expressed from the same gene without
reference to a sequenced genome.  Compared against competing tools
for the same purpose \cite{Burke99,Slater00,Huang99,Parkinson02,Kalyanaraman03,Malde03,
  Ptitsyn05,Hazelhurst08a,Picardi09}, {\it Peace} proves to be both more sensitive and
more robust to sequencing error without sacrificing runtime.

In short, {\sc Peace} is computational tool for the {\it ab initio}
clustering of transcript fragments by gene association, applicable to both NGS and
traditional Sanger Sequencing technologies.  Available through the
\href{http://www.peace-tools.org}{www.peace-tools.org} website, the
{\sc Peace} GUI allows the user to both easily install (locally or
remotely) and run the clustering engine, as well as enabling
transparent parallel processing and providing various tools for result
analysis.

\enlargethispage{-65.1pt}

\section{{\sc Peace}: Installation and Use}

The {\sc Peace} GUI can be launched by downloading and executing the
Jar file available on the {\sc Peace} website
(\href{http://www.peace-tools.org}{www.peace-tools.org}) to any
machine running the standard Java Virtual Machine (JVM).  Once running,
the user can employ the GUI to install the computational tool and perform
a clustering of a data file in the FASTA format, view an initial
analysis of the clusters, and produce files containing subsets of the
clusters as input to assembly tools such as {\sc Cap3} \cite{Huang99}.
A typical (first) use of {\sc Peace} generally
requires the following steps.

\noindent {\bf Tool Installation:} (First use only.)  To install the
{\sc Peace} clustering onto a local or remote machine, the
user selects from within the GUI the appropriate menu tab
(Figure~1(a)), which then starts an install wizard that will prompt
for the appropriate information.  Figure~1(b) illustrates the request
for server information; the user has chosen to install the {\sc Peace}
computational tool on a remote machine and is providing the necessary
connection information.  Server information is persistent between GUI
sessions, giving the user access to {\sc Peace} on the target machine
as needed.

\noindent {\bf Job Processing:} After importing the target sequence file into
the GUI, the user starts a new job by following the wizard menus.
Figure~1(c) illustrates the process of specifying the number of
processors available (if running on a machine supporting the OpenMPI
protocol -- which will be determined during job installation).  Once
executed, the GUI will manage the job thread, alert the user when the
job is completed (or when the user next runs the GUI after completion),
and copy the final results back to the local machine if necessary.

\noindent {\bf Result Analysis:} Once the resulting clusters have been
computed, the user has several options for analysis:
\begin{itemize}
\item {\bf Export:} The user can export the contents of one or more
  clusters into a FASTA format file, obtaining a subset of the original target file
  containing the sequences corresponding to the selected clusters
  ready for processing by an assembly tool (e.g. {\sc Cap3} \cite{Huang99}).
\item {\bf View Clustering}: The user may view a list of clusters,
  expanding selected clusters to a list of all individual sequences (illustrated
  in Figure~1(d)).
\item {\bf Classified Summary Graph:} The user may view a distribution
  of cluster sizes.  Further, the user can set up a {\it classifier},
  associating certain patterns with specific colors.  These patterns
  ware matched against the fragment header information from the
  original FASTA file, allowing the overlay of a colored cluster size
  distributions.  For example, if the sequence names contain unique
  string patterns denoting different cDNA libraries, the classifier
  can help the user to determine and visualize the differential expression profiles
  of different libraries for a given cluster.  The method of setting up these classifiers, and the
  resulting histogram, is illustrated in Figure~1(e).
\end{itemize}
Extensive documentation for the tool has been posted on the {\sc
  Peace} website, as well as links to several tutorial videos demonstrating
{\sc Peace} use and capabilities.

\section{Methods}

The clustering performed by {\sc Peace} is based on the use of minimum
spanning trees (MSTs), known to be an effective approach for narrow
band single linkage clustering \cite{Jain99,Wan08}. Edge weights are
assigned using the $d^2$ distance measure \cite{Hide94} to represent
edge weights.  Prim's algorithm \cite{Prim57} then allows us to
efficiently calculate an MST from which we can infer a high-quality
clustering solution.

The $d^2$ distance measure used to assign edge weights is an
alignment-free measurement of sequence distance that can be calculated
significantly faster than a Smith-Waterman alignment \cite{Hide94}.
$d^2$ works by comparing the frequency of words (strings of a fixed
length) appearing in a limited region of each string.  Fragments
overlapping by a sufficient length will share neighborhoods of enough
similarity to ensure a small distance even in the presence of a
moderate number of base errors.

Fragment input is modeled as a weighted, undirected graph: the fragments
are represented as nodes, with $d^2$ sequence distances assigned to
the connecting edges as weights.  Conceptually, we want to remove each
edge exceeding a threshold score from the complete graph and define
our partitions by the remaining connected components of the graph.  An
edge with a large weight connects fragments which are likely unrelated;
once such edges are removed the components define a series of
overlaps.  Those fragments that can still be connected by some path
correspond to the same gene.  However, such an approach requires the
calculation of all edge weights.  Such a task is infeasible both in terms of
runtime and memory usage for the data set sizes we expect to process.

{\sc Peace} approaches the problem by generating a minimum spanning tree of the
described graph, then removing edges exceeding our threshold.  By
using Prim's algorithm we are able to calculate edge weights
on-the-fly (reducing memory requirements) and can skip the
calculation of a majority of edge distances using the $u/v$ and $t/v$
filtering heuristics employed in {\sc wcd} \cite{Hazelhurst08a}.
These heuristics allow us to quickly dismiss many of the edges as too
large without the need to apply the full $d^2$ algorithm (see
Sections~A and B of the Supplementary Materials for more details).

\section{Results}

{\sc Peace} has been test on both simulated and real data from NGS and
Sanger Sequencing technologies, comparing results against those
produced by the {\sc wcd} clustering tool \cite{Hazelhurst08a} and the
{\sc Cap3} assembly tool \cite{Huang99} (which calculates a cluster as
an implicit, intermediate step).  For our simulation tests we used the
{\bf ESTSim} tool \cite{Hazelhurst03} to generate simulated transcript
of varying length fragments under different models of error
(Supplementary Materials, Section~C.1), generating the fragments from
the list of 100 zebra fish genes used in the {\sc wcd} testing
\cite{Hazelhurst08a}.  Tool parameters were taken to match, as closely
as possible, those used in the {\sc wcd} study (see supplementary
materials).  Our primary methods of quality assessment were {\it
  sensitivity} (the fraction of fragment pairs from the same gene that
were correctly clustered together) and {\it Type 1 error} (the
fraction of genes that were divided between clusters)
\cite{Wang04,Hazelhurst08a}; other measurements are also discussed in
the supplementary materials.  In Figure~2(left) we see a significant
improvement of {\sc Peace} (blue) over {\sc wcd} (green) and {\sc
  Cap3} (black) in its sensitivity to sequencing errors, while in
Figure~2(right) we see a comparable improvement in Type 1 error.  In
Figure~3 we look at the number of singleton clusters (fragments not
joined to any cluster), which will not occur in our simulated sets; we
again see significant improvement in {\sc Peace}.  The sequential
runtime of {\sc Peace} is slightly improved over that of {\sc wcd}
(see Supplementary Materials, Figure~S4).

\begin{figure*}[b]
\centerline{
\includegraphics[trim=0cm 3cm 0cm 1cm, clip, scale=0.9]{screen.d/screen.pdf} 
}
\label{screen}
\caption{Screenshots of the {\sc Peace} GUI during execution, including (a)
GUI Welcome and server installation menu; (b) setup wizard for
installing the computational tool on a remote server; (c) execution
wizard for starting a selected job to be executed in parallel mode;
(d) basic cluster output; and (e) histogram view of cluster results
and classifier editor for setting up differential expression profiles.}
\end{figure*}

\begin{figure*}[h]
\centerline{
\begin{minipage}[b]{2in}
\includegraphics[scale=0.5]{pics.d/SeT1.pdf} 
\end{minipage}{2in}
}
\label{SeT1}
\caption{Comparisons of Sensitivity and Type 1 error, based on the
  average over 30 simulated fragment sets derived from 100 zebra fish genes
  (see Supplementary Materials, Section~\href{sim_results}, for more
  details).   Blue/Solid = {\sc Peace}, Green/Dash =
  {\sc wcd}, Black/Dot-Dash = {\sc Cap3}; vertical tics = 95\% confidence
  intervals on estimates.  Intervals are not presented for Type 1
  error due to the extremely small variance.}
\end{figure*}

\begin{figure*}[b]
\begin{minipage}{2in}
\centerline{
\includegraphics[scale=0.35]{pics.d/singletons.pdf}
}
\end{minipage}
\label{singletons}
\caption{Average number of fragments flagged as singletons by each tool
  when run on the simulated sequences; correct answer in all cases is
  zero.  Blue/Solid = {\sc Peace}, Green/Dash = {\sc wcd},
  Black/Dot-Dash = {\sc Cap3}.}
\end{figure*}

In applying the tools to real data, we started with a set of
approximately $190,000$ clean Sanger ESTs derived from the {\it
  Chlamydomonas reinhardtii} genome \cite{Liang2008}.  To compute
sensitivity we used the {\it gmap} tool \cite{Wu05} to map the
fragment set to the genome, taking this as our reference clustering.
We see slight improvements in {\sc Peace} over {\sc wcd} in both
sensitivity and Type 1 error (both significantly outperforming {\sc
  Cap3}).  Using the mouse data set used in Hazelhurst {\it et al.}
\cite{Hazelhurst08a} {\sc Peace} again shows slight improvements in
sensitivity and Type 1 error rates, with an 18\% speedup for {\sc
  Peace} (see Supplementary Materials, Section~D, for more details).

\section{Conclusions}

Here we have presented {\sc Peace}, a stand alone tool for the
high-throughput clustering of transcript fragments, managed through a
user friendly GUI that enables both local and remote installation and
execution in sequential or parallel mode.  Based on a novel algorithm
for the clustering of the fragments by gene association, {\sc Peace}
shows significant improvement in sensitivity, without sacrificing runtime,
over the competing {\sc WCD} tool \cite{Hazelhurst08a}, and shows an
order of magnitude in improvement over the clustering preformed in the
course of assembly by the {\sc Cap3} tool \cite{Huang99}.

As a clustering tool based on sequence distance, {\sc Peace} faces
certain inherent limitations. For example, {\sc Peace} cannot handle
duplicate genes; like {\sc WCD}, it is unable to separate clusters
corresponding to genes with a greater than 88\% similarity.
Similarly, natural biological effects such as the trans-splicing of
transcripts, or effects from poorly cleaned transcript data (e.g. the
failure to remove sequencing adapters or post-transcriptional
poly(A)/(T) tails), and low-complexity repeats can cause similar
effects in these clustering tools.  The problems can be handled
through the application of the assembler, and the ability to apply any
assembler to small cluster (as opposed to the data set as a whole)
results in a significant reduction in overall assembly time.

\section{Acknowledgements}

Dr. Karro was funded under a PhRMA Foundation Informatics Research
Starters Grant while conducting this research.  We would also like to
acknowledge Iddo Friedberg, David Woods, Elizabeth Bikun, Jens Mueller and David
Scoville at Miami University for their help with this project.

\vspace{3mm}
%Bib TeX

\bibliography{peace.bib}

\end{document}

% LocalWords:  arallel nalysis lustering ngine Rao Moler Ozden Zhang Liang ESTs
% LocalWords:  Karro XXXXX transcriptome pre isoforms SNPs unsequenced runtime
% LocalWords:  chlamydomondomonas ESTate PaCE xsact CLU WCD Hazelhurst et al th
% LocalWords:  CHUN nalyzer Prim's MSTs subgraph MPI SourceCode ESTSim runtimes
% LocalWords:  ests Chlamydomonas Reihartdii gmap CDNA transcriptomes cDNA www
% LocalWords:  reinhartdii conifergdb DNAc bp wcd reinhardtii PhRMA Iddo fasta
% LocalWords:  Friedberg Scoville OpenMPI initio multi Screenshots nvironment
% LocalWords:  ssembly xpression GenBank dbEST pyrosequencing ESTsim sim mRNAs
% LocalWords:  NGS polyA