### 2010-04-18 papers/NAR_web_server_2009/paper.tex
Modifications to introduction of response to reviewers

\documentclass[a4,center,fleqn]{NAR}
\usepackage[pdftex,
colorlinks=true, %makes the links show up in color, rather than in a box
citecolor=black, %color of the in-text citation numbers
urlcolor=blue %color of the url links
]{hyperref}
\usepackage{wrapfig}
\bibliographystyle{unsrt}

% Enter dates of publication
\copyrightyear{}
\pubdate{}
\pubyear{}
\jvolume{}
\jissue{}

\usepackage[usenames,dvipsnames]{color}
\definecolor{red}{rgb}{1,0,0}
\newcommand{\mc}[1]{\textcolor{red}{#1}}

% \articlesubtype{This is the article type (optional)}

% Some shortcuts to make life easier when writing latex
\newcommand{\peace} {{\small PEACE}}
\newcommand{\wcd} {{\small WCD}}
\newcommand{\capthree} {{\small Cap3}}
\newcommand{\easycluster} {{\small EasyCluster}}

\begin{document}

\title{PEACE: {\underline P}arallel {\underline E}nvironment for {\underline A}ssembly
  and {\underline C}lustering of Gene {\underline E}xpression}

\author{D.M. Rao\,$^{1}$, J.C. Moler\,$^{1}$, M. Ozden\,$^1$, Y. Zhang\,$^{1}$,
  C. Liang$^{1,2*}$\, and J.E. Karro\,$^{1,3}$\footnote{to whom
    correspondence should be addressed}}

\address{$^1$ Department of Computer Science and Software Engineering, \\
  $^2$ Department of Botany, \\
  $^3$ and Department of Microbiology, Miami University, Oxford, Ohio,
  USA}

\history{Received Feb. 13, 2010}

\maketitle

\begin{abstract}
  We present \peace, a stand alone tool for the high-throughput {\it
    ab initio} clustering of transcript fragment sequences produced by
  Next Generation or Sanger Sequencing technologies, freely available
  from \href{http://www.peace-tools.org}{www.peace-tools.org}.
  Installed and run through an easily downloaded GUI, \peace\/ can
  process large data sets of transcript fragments of length 50 bases
  or greater, grouping the fragments by gene association and achieving
  a greater sensitivity (without sacrificing runtime) than the
  competing clustering tools.  Once clustered, the user can employ the
  GUI to collect statistics and single out specific clusters for more
  comprehensive study.  \peace\/, when used as a ``pre-assembly''
  tool, enables any assembly tool to work with individual small
  clusters of closely related sequences.  For a 5\% sequencing error
  rate, we see a 52\% improvement in sensitivity over the leading
  cluster tool, \wcd\/, and 184\% improvement in sensitivity over the
  assembly tool \capthree.  When applied to a 361 Mb set of
  Arabidopsis transcriptome fragments, we see a 40\% speedup over
  \wcd.
\end{abstract}

\section{Introduction}

Understanding an organism's transcriptome, the set of (spliced)
transcripts expressed by genes of the organism, is a vital step in
understanding the full functional and organizational role of the
genome in the life cycle of any eukaryote.  Studying the transcriptome
has led to gene discovery, provided information on splice variants,  
and helped shed light on the biological processes both controlling and
controlled by the genome \cite{Nagaraj07}.  However, to access those
transcripts, we must deal with the fragmented data produced by
both Next Generation and traditional Sanger sequencing technology.

In the past, access to a transcriptome sequence was primarily
through the use of Expressed Sequence Tags (ESTs), single-pass cDNA
sequences derived from transcribed mRNAs and sequenced by Sanger
Sequencing technology.  More recently, Next Generation Sequencing
(NGS) technology has begun to rapidly replace Sanger Sequencing,
allowing for more comprehensive coverage of the transcriptome.  For
example, ESTs now being added to the GenBank dbEST are increasingly
the product of NGS technologies such as 454 pyrosequencing, which
enables the sequencing of novel and rare transcripts at a considerably
higher rate of coverage than Sanger Sequencing
\cite{Cheung2006,Emrich2007}.  From a computational perspective, this
is a mixed blessing: while NGS provides immense quantities of new
information, it also provides immensely larger data sets -- and thus a
need for faster, more efficient analysis algorithms.

\enlargethispage{-65.1pt}

Given a set of transcript fragments sampled from across the genome, a
necessary first step of the set's analysis is that of clustering: separating
the fragments according to the transcript from which they were
derived.  Frequently performed implicitly by assembly tools,
clustering the data as a ``pre-assembly'' step has a number of
advantages.  Most significantly: performing this step will
allow the application of the assembly tool to individual clusters --
saving significant amounts of time \cite{Hazelhurst08a}.

However, clustering is a computationally challenging problem.  Even
with the smaller number of ESTs produced using Sanger Sequencing, the
runtime and memory requirements to cluster on the basis of pair-wise
sequence alignments make such an approach infeasible in practice.  The
much larger data-set size produced by NGS technologies exacerbates
this problem.  To deal with this, \peace\/ combines our own version of
the $d^2$ alignment-free sequence distance function \cite{Hide94} and
the concept of a {\it minimum spanning tree} \cite{Prim57} to quickly
and accurately find clusters of ESTs expressed from the same gene
without reference to a sequenced genome.  Compared against \wcd, the
leading clustering tool in the literature \cite{Hazelhurst08a}, as well as
other tools designed for the same purpose 
\cite{Burke99,Slater00,Huang99,Parkinson02,Kalyanaraman03,Malde03,
  Ptitsyn05,Hazelhurst08a,Picardi09}, \peace\/ proves to be both more
sensitive and more robust to sequencing error without sacrificing
runtime.  Nor are any of these tools designed for the ease of installation
and use that \peace provides.

In short, \peace\/ is a computational tool for the {\it ab initio}
clustering of transcript fragments by gene association, applicable to
both NGS and traditional Sanger Sequencing technologies.  Available
through the \href{http://www.peace-tools.org}{www.peace-tools.org}
website, the \peace\/ GUI allows the user to both easily install
(locally or remotely) and run the clustering engine, as well as
enabling transparent parallel processing and providing various tools
for result analysis.

\section{\peace\/: Installation and Use}

The \peace\/ GUI can be launched by downloading and executing the
Jar file available on the \peace\/ website
(\href{http://www.peace-tools.org}{www.peace-tools.org}) to any
machine running the standard Java Virtual Machine (JVM).  Once
running, the user can employ the GUI to install the clustering engine
and perform a clustering of a data file in {\sc fasta} format, view an
initial analysis of the clusters, and produce files containing subsets
of the clusters as input to assembly tools such as \capthree\/ 
\cite{Huang99}.  A typical (first) use of \peace\/ must be
performed in the following manner (see Figure~\ref{fig:workflow}):

\begin{figure}
  \centerline{\includegraphics[width=3in]{pics.d/workflow.eps}}
  \caption{Overview of the procedure for clustering and analysis using
    PEACE.}\label{fig:workflow}
\end{figure}

\noindent {\bf Tool Installation:} (First use only.) To install the
\peace\/ clustering engine onto a local or remote machine, the user selects
from within the GUI the appropriate menu tab (Figure~\ref{screen}(a)),
which then starts an install wizard that will prompt for the
appropriate information.  Figure~\ref{screen}(b) illustrates the
request for server information; the user has chosen to install the
\peace\/ computational tool on a remote machine and is providing
the necessary connection information.  Server information is
persistent between GUI sessions, giving the user access to \peace\/
on the target machine as needed.

\begin{figure*}
  \begin{minipage}{3in}
    \includegraphics[width=3in]{install_page_big.eps}
    \centerline{\small{(a)}}
  \end{minipage}
  \begin{minipage}{2in}
    \vspace{0.47in}
    \includegraphics[width=2in]{add_server_big.eps}
    \centerline{\small{(b)}}
  \end{minipage}
  \begin{minipage}{2in}
    \vspace{0.47in}
    \includegraphics[width=2in]{processor_assignment_big.eps}
    \centerline{\small{(c)}}
  \end{minipage}

  \begin{minipage}{3in}
    \includegraphics[width=3in]{cluster_list_big.eps}
    \centerline{\small{(d)}}
  \end{minipage}
  \begin{minipage}{1in}
    \hspace{1in}
  \end{minipage}
  \begin{minipage}{3in}
    \includegraphics[width=3in]{classifier_big.eps}
    \centerline{\small{(e)}}
  \end{minipage}

  \caption{Screenshots of the \peace\/ GUI during execution, including (a)
    GUI Welcome and server installation menu; (b) setup wizard for
    installing the computational tool on a remote server; (c) execution
    wizard for starting a selected job to be executed in parallel mode;
    (d) basic cluster output; and (e) histogram view of cluster results
    and classifier editor for setting up differential expression profiles.}\label{screen}
\end{figure*}

\noindent {\bf Job Processing:} After importing the target sequence
file into the GUI, the user starts a new job by following the wizard
menus.  Figure~\ref{screen}(c) illustrates the process of specifying
the number of processors available (if running on a machine supporting
the OpenMPI protocol -- which will be determined during job
installation).  Once executed, the GUI will manage the job thread,
alert the user when the job is completed (or when the user next runs
the GUI after completion), and copy the final results back to the
local machine if necessary.

\noindent {\bf Result Analysis:} Once the resulting clusters have been
computed, the user has several options for analysis:
\begin{itemize}
\item {\bf Export:} The user can export the contents of one or more
  clusters into a FASTA format file, obtaining a subset of the
  original target file containing the sequences corresponding to the
  selected clusters ready for processing by an assembly tool
  (e.g. \capthree\/ \cite{Huang99}).
\item {\bf View Clustering}: The user may view a list of clusters,
  expanding selected clusters to a list of all individual sequences
  (illustrated in Figure~\ref{screen}(d)).
\item {\bf Classified Summary Graph:} The user may view a distribution
  of cluster sizes.  Further, the user can set up a {\it classifier},
  associating certain patterns with specific colors.  These patterns
  ware matched against the fragment header information from the
  original FASTA file, allowing the overlay of a colored cluster size
  distributions.  For example, if the sequence names contain unique
  string patterns denoting different cDNA libraries, the classifier
  can help the user to determine and visualize the differential
  expression profiles of different libraries for a given cluster.  The
  method of setting up these classifiers, and the resulting histogram,
  is illustrated in Figure~\ref{screen}(e).
\end{itemize}
Extensive documentation for the tool has been posted on the \peace\/ website, as well as links to several tutorial videos
demonstrating \peace\/ use and capabilities.

\section{Methods}

The clustering performed by \peace\/ is based on the use of minimum
spanning trees (MSTs), known to be an effective approach for narrow
band single linkage clustering \cite{Jain99,Wan08}.  Using a graph
structure to model the fragment relationships and the $d^2$ distance
measure to assign edge weights \cite{Hide94}, we can employ Prim's
algorithm \cite{Prim57} to efficiently calculate an MST from which we
can infer a high-quality clustering solution.

The $d^2$ distance measure used to assign edge weights is an
alignment-free measurement of sequence distance that can be calculated
significantly faster than a Smith-Waterman alignment \cite{Hide94}.
$d^2$ works by comparing the frequency of words (strings of a fixed
length) appearing in a limited region of each string.  Fragments
overlapping by a sufficient length will share neighborhoods of enough
similarity to ensure a small distance even in the presence of a
moderate number of base errors.  In practice we employ our own
variation of $d^2$, the {\it two-pass $d^2$ algorithm}, which
heuristically searches for a neighborhood of maximum similarity and
then finds the $d^2$ score based on that neighborhood (see
Supplementary Materials for details).

Fragment input is modeled as a weighted, undirected graph: the
fragments are represented as nodes, with $d^2$ sequence distances
assigned to the connecting edges as weights.  Conceptually, we want to
remove each edge exceeding a threshold score from the complete graph
and define our partitions by the remaining connected components.  An
edge with a large weight connects fragments which are likely
unrelated; once such edges are removed the components define a series
of overlaps.  Those fragments that can still be connected by some path
correspond to the same gene.  However, such an approach requires the
calculation of all edge weights.  That task is infeasible both in terms
of runtime and memory usage for the data set sizes we expect to process.

\peace\/ approaches the problem by generating a minimum spanning tree of the
described graph, then removing edges exceeding our threshold.  By
using Prim's algorithm we are able to calculate edge weights
on-the-fly (reducing memory requirements) and can skip the
calculation of a majority of edge distances using the $u/v$ and $t/v$
filtering heuristics employed in \wcd\/ \cite{Hazelhurst08a}.
These heuristics allow us to quickly dismiss many of the edges as too
large without the need to apply the full $d^2$ algorithm (see
Sections~A and B of the Supplementary Materials for more details).

\section{Results}

\peace\/ has been tested on both simulated and real data from NGS and
Sanger Sequencing technologies, comparing results against those
produced by the \wcd\/ clustering tool \mc{(version
  0.5.1)}\cite{Hazelhurst08a} and the \capthree\/ assembly tool
\cite{Huang99} (the latter of which implicitly calculates a clustering
in the process of assembly).  For our simulation tests we used the
{\bf ESTSim} tool \cite{Hazelhurst03} to generate simulated transcript
fragments of varying length under different models of error
(Supplementary Materials, Section~C.1), generating the fragments from
the list of 100 zebra fish genes used in the \wcd\/ testing
\cite{Hazelhurst08a}.  Tool parameters were taken to match, as closely
as possible, those used in the \wcd\/ study (see Supplementary
Materials).  \mc{The most important method of quality assessment is
  {\it sensitivity} (the fraction of fragment pairs from the same gene
  that were correctly clustered together).  But we also look at the
  {\it Jaccard Index} (which balances sensitivity with the number of
  false positives by looking at the ratio of true positives to the sum
  of true positives, false positives, and true negitives), {\it Type 1
    error} (the fraction of genes that were divided between clusters),
  and {\it Type 2 error} (the fraction of clusters containing two or
  more genes).  Specificity was not an issue in simulated tests, as
  both \peace\/ and \wcd\/ were completely successful in separating
  unrelated ESTs (save when dealing with recently duplicated genes).
  Other measurements are also discussed in the supplementary
  materials.  In Figure~\ref{SeJiT1T2} (left) we see these four tests,...}

% , reflecting tests done on
%   simulated Sanger EST sets, we see a significant improvement of
%   \peace\/ (blue) over \wcd\/ (green) and \capthree\/ (black) in its
%   sensitivity to sequencing errors, while in Figure~\ref{SeT1} (right)
%   we see a comparable improvement in Type 1 error.  In
%   Figure~\ref{singletons} we look at the number of singleton clusters
%   (fragments not joined to any cluster), which should not occur in our
%   simulated sets.  We again see significant improvement in \peace.
%   The sequential runtime of \peace\/ is slightly improved over that of
%   \wcd\/ (see Supplementary Materials, Figure~S4).

\begin{figure}
  \centerline{\includegraphics[width=3.35in]{pics.d/SeJiT1T2.40.eps}}
  \caption{Comparisons of Sensitivity and Type 1 error, based on the
    average over 30 simulated Sanger Sequence ESTs sets derived from 100 zebra
    fish genes  (see
    Supplementary Materials, Section C, for more details).  Blue/Solid
    = \peace, Green/Dash = \wcd, Black/Dot-Dash = \capthree; vertical
    tics = 95\% confidence intervals on estimates.  Intervals are not
    presented for Type 1 error due to the extremely small
    variance.}\label{SeJiT1T2}
\end{figure}

\mc{In applying the tools to real data Sanger data, we used the Human
Benchmark Dataset used to test \easycluster\/ \cite{Picardi09}, and
the A076941.fa {\it Arabidopsis thaliana} dataset used to test WCD
\cite{Hazelhurst08,Hazelhurst08a}.  For the first, we find an almost
exact match in results in quality between PEACE and WCD (both
out-preforming Cap3), but a significant gain in runtime for PEACE over
WCD, as shown in Tabe\ref{XXX}}.  For the A076941.fa...}

% we started with a set of
% approximately $190,000$ clean Sanger ESTs derived from the {\it
%   Chlamydomonas reinhardtii} genome \cite{Liang2008}.  To compute
% sensitivity we used the {\it gmap} tool \cite{Wu05} to map the
% fragment set to the genome, taking this as our reference clustering.
% We see slight improvements in \peace\/ over \wcd\/ in both sensitivity
% and Type 1 error (both significantly outperforming \capthree).  Using
% the \mc{Arabidopsis} data set used in Hazelhurst {\it et al.}
% \cite{Hazelhurst08a} \peace\/ again shows slight improvements in
% sensitivity and Type 1 error rates, with an 18\% speedup for \peace\/ 
% (see Supplementary Materials, Section~D, for more details).

We were unable to run \wcd\/ on short-read sequences, as the
distributed implementation is unable to handle such data without an
adjustment to the code.  Hence, while \peace\/ was run on both
simulated and real NGS fragment sets, there was no cluster tool to
which the results could be compared.  \peace\/ did well, achieving a
sensitivity of 0.75 when applied to a simulated short-read set
subjected to a sequencing error rate of 2\% (with comparable {\it
  Jaccard Index} values reflecting both sensitivity {\it and}
specificity \cite{Kalyanaraman03,Hazelhurst08a}).  Runtime was
approximately 17.6 minutes to process the 7.9 Mb of data when run in
sequential mode on a \mc{3.2 GHz Intel Xeon EM64T CPU with a 2 MB
  cache and 800 MHz font side bus}.  When applied to a 64 Mb set of
165,490 {\it
 Chlamydomonas reinhardtii} transcript fragments
sequenced by 454 pyrosequencing technology (ranging in size from 50
bases to several hundred), \peace\/ completed the job in 49 minutes
using 30 of the Xeon processors.

\section{Conclusions}

Here we have presented \peace, a stand alone tool for the
high-throughput clustering of transcript fragments capable of dealing
with sequences as short as 50 bases.  \peace\/, available at
\href{http://www.peace-tools.org}{www.peace-tools.org}, is open-source
and managed through a user friendly GUI that enables both local and
remote installation and execution in sequential or parallel mode.
Based on a novel algorithm for the clustering of the fragments by gene
association, \peace\/ shows significant improvement in sensitivity,
without sacrificing runtime, over the competing \wcd\/ tool
\cite{Hazelhurst08a}, and shows an order of magnitude in improvement
over the clustering performed in the course of assembly by the
\capthree\/ tool \cite{Huang99}.

As a clustering tool based on sequence distance, \peace\/ faces
certain inherent limitations. For example, \peace\/ cannot handle
duplicate genes; like \wcd\/, it is unable to separate clusters
corresponding to genes with a greater than 88\% similarity.
Similarly, other natural biological effects (e.g. the trans-splicing of
transcripts), effects from poorly cleaned transcript data (e.g. the
failure to remove sequencing adapters or post-transcriptional
poly(A)/(T) tails), and the presence of low-complexity repeats can
cause similar effects in these clustering tools.  The problems can be
handled through the application of the assembler, and the ability to
apply any assembler to small cluster (as opposed to the data set as a
whole) results in a significant reduction in overall assembly time.

\section{Acknowledgements}

Dr. Karro was funded under a PhRMA Foundation Informatics Research
Starters Grant while conducting this research.  We would also like to
acknowledge Iddo Friedberg, David Woods, Elizabeth Bikun, Jens Mueller and David
Scoville at Miami University for their help with this project.

\vspace{3mm}
% Bib TeX

\bibliography{peace.bib}

\end{document}

% LocalWords:  arallel nalysis lustering ngine Rao Moler Ozden Zhang Liang ESTs
% LocalWords:  Karro XXXXX transcriptome pre isoforms SNPs unsequenced runtime
% LocalWords:  chlamydomondomonas ESTate PaCE xsact CLU WCD Hazelhurst et al th
% LocalWords:  CHUN nalyzer Prim's MSTs subgraph MPI SourceCode ESTSim runtimes
% LocalWords:  ests Chlamydomonas Reihartdii gmap CDNA transcriptomes cDNA www
% LocalWords:  reinhartdii conifergdb DNAc bp wcd reinhardtii PhRMA Iddo fasta
% LocalWords:  Friedberg Scoville OpenMPI initio multi Screenshots nvironment
% LocalWords:  ssembly xpression GenBank dbEST pyrosequencing ESTsim sim mRNAs
% LocalWords:  NGS polyA set's GHz Xeon Bikun