### 2010-04-21 peace/papers/NAR_web_server_2009/paper.tex
Final version of manuscript

\documentclass[a4,center,fleqn]{NAR}
\usepackage[pdftex,
colorlinks=true, %makes the links show up in color, rather than in a box
citecolor=black, %color of the in-text citation numbers
urlcolor=blue %color of the url links
]{hyperref}
\usepackage{wrapfig}
\bibliographystyle{unsrt}

% Enter dates of publication
\copyrightyear{}
\pubdate{}
\pubyear{}
\jvolume{}
\jissue{}

\usepackage[usenames,dvipsnames]{color}
\definecolor{red}{rgb}{1,0,0}
\newcommand{\mc}[1]{\textcolor{red}{#1}}

% \articlesubtype{This is the article type (optional)}

% Some shortcuts to make life easier when writing latex
\newcommand{\peace} {{\small PEACE}}
\newcommand{\wcd} {{\small WCD}}
\newcommand{\capthree} {{\small Cap3}}
\newcommand{\easycluster} {{\small EasyCluster}}
\newcommand{\metasim} {{\small MetaSim}}

\begin{document}

\title{PEACE: {\underline P}arallel {\underline E}nvironment for {\underline A}ssembly
  and {\underline C}lustering of Gene {\underline E}xpression}

\author{D.M. Rao\,$^{1}$, J.C. Moler\,$^{1}$, M. Ozden\,$^1$, Y. Zhang\,$^{1}$,
  C. Liang$^{1,2*}$\, and J.E. Karro\,$^{1,3}$\footnote{to whom
    correspondence should be addressed}}

\address{$^1$ Department of Computer Science and Software Engineering, \\
  $^2$ Department of Botany, \\
  $^3$ and Department of Microbiology, Miami University, Oxford, Ohio,
  USA}

\history{Received Feb. 13, 2010 \\ Resubmitted April 20, 2010}

\maketitle

\begin{abstract}
  \mc{We present \peace, a stand alone tool for high-throughput {\it
      ab initio} clustering of transcript fragment sequences produced
    by Next Generation or Sanger Sequencing technologies.  It is
    freely available from
    \href{http://www.peace-tools.org}{www.peace-tools.org}.  Installed
    and managed through a download-able user-friendly GUI, \peace\/
    can process large data sets of transcript fragments of length 50
    bases or greater, grouping the fragments by gene associations with
    a sensitivity compatible to leading clustering tools.  Once
    clustered, the user can employ the GUI's analysis functions,
    facilitating the easy collection of statistics and allowing them
    to single out specific clusters for more comprehensive study or
    assembly.  Using a novel {\it minimum spanning tree} based
    clustering method, \peace\/ is the equal of other leading tools in
    the literature in quality with an interface making it
    accessible to any user.  It produces results of quality virtually
    identical to those of the \wcd\/ tool when applied to Sanger
    sequences, significantly improved results over \wcd\/ when applied
    the products of Next Generation Sequencing Technology, and
    significantly improved results over \capthree\/ in both cases.  In
    short, \peace\/ provides an intuitive graphical user interface and
    a feature-rich, parallel clustering engine that proves to be a
    valuable addition to the leading cDNA clustering tools.}

\end{abstract}

\section{Introduction}

Understanding an organism's transcriptome, the set of (spliced)
transcripts expressed by genes of the organism, is a vital step in
understanding the full functional and organizational role of the
genome in the life cycle of any eukaryote.  Studying the transcriptome
has led to gene discovery, provided information on splice variants,  
and helped shed light on the biological processes both controlling and
controlled by the genome \cite{Nagaraj07}.  However, to access those
transcripts, we must deal with the fragmented data produced by
both Next Generation and traditional Sanger sequencing technology.

In the past, access to a transcriptome sequence was primarily
through the use of Expressed Sequence Tags (ESTs), single-pass cDNA
sequences derived from transcribed mRNAs and sequenced by Sanger
Sequencing technology.  More recently, Next Generation Sequencing
(NGS) technology has begun to rapidly replace Sanger Sequencing,
allowing for more comprehensive coverage of the transcriptome.  For
example, ESTs now being added to the GenBank dbEST are increasingly
the product of NGS technologies such as 454 pyrosequencing, which
enables the sequencing of novel and rare transcripts at a considerably
higher rate of coverage than Sanger Sequencing
\cite{Cheung2006,Emrich2007}.  From a computational perspective, this
is a mixed blessing: while NGS provides immense quantities of new
information, it also provides immensely larger data sets -- and thus a
need for fast, efficient analysis algorithms.

\enlargethispage{-65.1pt}

Given a set of transcript fragments sampled from across the genome, a
necessary first step of the set's analysis is that of clustering: separating
the fragments according to the transcript from which they were
derived.  Frequently performed implicitly by assembly tools,
clustering the data as a ``pre-assembly'' step has a number of
advantages.  Most significantly: performing this step will
allow the application of the assembly tool to individual clusters --
saving significant amounts of time \cite{Hazelhurst08a}.

However, clustering is a computationally challenging problem.  Even
with the smaller number of ESTs produced using Sanger Sequencing, the
runtime and memory requirements to cluster on the basis of pair-wise
sequence alignments make such an approach infeasible in practice.  The
much larger data-set size produced by NGS technologies exacerbates
this problem.  To deal with this, \peace\/ combines our own version of
the $d^2$ alignment-free sequence distance function \cite{Hide94} and
the concept of a {\it minimum spanning tree} \cite{Prim57} to quickly
and accurately find clusters of ESTs expressed from the same gene
without reference to a sequenced genome.  \mc{Compared against
  clustering tool in the literature
  \cite{Burke99,Slater00,Huang99,Parkinson02,Kalyanaraman03,Malde03,
    Ptitsyn05,Hazelhurst08a,Picardi09}, \peace\/ proves to be of
  equivalent quality to the \wcd\/ tool \cite{Hazelhurst08a}, more
  sensitive and robust than the other tools, and more user friendly
  than any published clustering tool.  Nor are any of} \mc{these tools
  designed for the ease of installation and use that \peace\/
  provides.}

In short, \peace\/ is a computational tool for the {\it ab initio}
clustering of transcript fragments by gene association, applicable to
both NGS and traditional Sanger Sequencing technologies.  Available
through the \href{http://www.peace-tools.org}{www.peace-tools.org}
website, the \peace\/ GUI allows the user to both easily install
(locally or remotely) and run the clustering engine, as well as
enabling transparent parallel processing and providing various tools
for result analysis.

\section{\peace: Installation and Use}

The \peace\/ GUI can be launched by downloading and executing the
Jar file available on the \peace\/ website
(\href{http://www.peace-tools.org}{www.peace-tools.org}) to any
machine running the standard Java Virtual Machine (JVM).  Once
running, the user can employ the GUI to install the clustering engine
and perform a clustering of a data file in {\sc fasta} format, view an
initial analysis of the clusters, and produce files containing subsets
of the clusters as input to assembly tools such as \capthree\/ 
\cite{Huang99}.  A typical (first) use of \peace\/ must be
performed in the following manner (see Figure~\ref{fig:workflow}):

\begin{figure}
  \centerline{\includegraphics[width=3in]{pics.d/workflow.png}}
  \caption{Overview of the procedure for clustering and analysis using
    \peace.}\label{fig:workflow}
\end{figure}

\noindent {\bf Tool Installation:} (First use only.) To install the
\peace\/ clustering engine onto a local or remote machine, the user selects
from within the GUI the appropriate menu tab (Figure~\ref{screen}(a)),
which then starts an install wizard that will prompt for the
appropriate information.  Figure~\ref{screen}(b) illustrates the
request for server information; the user has chosen to install the
\peace\/ computational tool on a remote machine and is providing
the necessary connection information.  Server information is
persistent between GUI sessions, giving the user access to \peace\/
on the target machine as needed.

\begin{figure*}
  \begin{minipage}{3in}
    \includegraphics[width=3in]{screen.d/install_page_big.png}
    \centerline{\small{(a)}}
  \end{minipage}
  \begin{minipage}{2in}
    \vspace{0.47in}
    \includegraphics[width=2in]{screen.d/add_server_big.png}
    \centerline{\small{(b)}}
  \end{minipage}
  \begin{minipage}{2in}
    \vspace{0.47in}
    \includegraphics[width=2in]{screen.d/processor_assignment_big.png}
    \centerline{\small{(c)}}
  \end{minipage}

  \begin{minipage}{3in}
    \includegraphics[width=3in]{screen.d/cluster_list_big.png}
    \centerline{\small{(d)}}
  \end{minipage}
  \begin{minipage}{1in}
    \hspace{1in}
  \end{minipage}
  \begin{minipage}{3in}
    \includegraphics[width=3in]{screen.d/classifier_big.png}
    \centerline{\small{(e)}}
  \end{minipage}

  \caption{Screenshots of the \peace\/ GUI during execution, including (a)
    GUI Welcome and server installation menu; (b) setup wizard for
    installing the computational tool on a remote server; (c) execution
    wizard for starting a selected job to be executed in parallel mode;
    (d) basic cluster output; and (e) histogram view of cluster results
    and classifier editor for setting up differential expression profiles.}\label{screen}
\end{figure*}

\noindent {\bf Job Processing:} After importing the target sequence
file into the GUI, the user starts a new job by following the wizard
menus.  Figure~\ref{screen}(c) illustrates the process of specifying
the number of processors available (if running on a machine supporting
the OpenMPI protocol -- which will be determined during job
installation).  Once executed, the GUI will manage the job thread,
alert the user when the job is completed (or when the user next runs
the GUI after completion), and copy the final results back to the
local machine if necessary.

\noindent {\bf Result Analysis:} Once the resulting clusters have been
computed, the user has several options for analysis:
\begin{itemize}
\item {\bf Export:} The user can export the contents of one or more
  clusters into a FASTA format file, obtaining a subset of the
  original target file containing the sequences corresponding to the
  selected clusters ready for processing by an assembly tool
  (e.g. \capthree\/ \cite{Huang99}).
\item {\bf View Clustering}: The user may view a list of clusters,
  expanding selected clusters to a list of all individual sequences
  (illustrated in Figure~\ref{screen}(d)).
\item {\bf Classified Summary Graph:} The user may view a distribution
  of cluster sizes.  Further, the user can set up a {\it classifier},
  associating certain patterns with specific colors.  These patterns
  ware matched against the fragment header information from the
  original FASTA file, allowing the overlay of a colored cluster size
  distributions.  For example, if the sequence names contain unique
  string patterns denoting different cDNA libraries, the classifier
  can help the user to determine and visualize the differential
  expression profiles of different libraries for a given cluster.  The
  method of setting up these classifiers, and the resulting histogram,
  is illustrated in Figure~\ref{screen}(e).
\end{itemize}
Extensive documentation for the tool has been posted on the \peace\/ website, as well as links to several tutorial videos
demonstrating \peace\/ use and capabilities.

\section{Methods}

The clustering performed by \peace\/ is based on the use of minimum
spanning trees (MSTs), known to be an effective approach for narrow
band single linkage clustering \cite{Jain99,Wan08}.  Using a graph
structure to model the fragment relationships and the $d^2$ distance
measure to assign edge weights \cite{Hide94}, we can employ Prim's
algorithm \cite{Prim57} to efficiently calculate an MST from which we
can infer a high-quality clustering solution.

The $d^2$ distance measure used to assign edge weights is an
alignment-free measurement of sequence distance that can be calculated
significantly faster than a Smith-Waterman alignment \cite{Hide94}.
$d^2$ works by comparing the frequency of words (strings of a fixed
length) appearing in a limited region of each string.  Fragments
overlapping by a sufficient length will share neighborhoods of enough
similarity to ensure a small distance even in the presence of a
moderate number of base errors.  In practice we employ our own
variation of $d^2$, the {\it two-pass $d^2$ algorithm}, which
heuristically searches for a neighborhood of maximum similarity and
then finds the $d^2$ score based on that neighborhood (see
Supplementary Materials for details).

Fragment input is modeled as a weighted, undirected graph: the
fragments are represented as nodes, with $d^2$ sequence distances
assigned to the connecting edges as weights.  Conceptually, we want to
remove each edge exceeding a threshold score from the complete graph
and define our partitions by the remaining connected components.  An
edge with a large weight connects fragments which are likely
unrelated; once such edges are removed the components define a series
of overlaps.  Those fragments that can still be connected by some path
correspond to the same gene.  However, such an approach requires the
calculation of all edge weights.  That task is infeasible both in terms
of runtime and memory usage for the data set sizes we expect to process.

\peace\/ approaches the problem by generating a minimum spanning tree of the
described graph, then removing edges exceeding our threshold.  By
using Prim's algorithm we are able to calculate edge weights
on-the-fly (reducing memory requirements) and can skip the
calculation of a majority of edge distances using the $u/v$ and $t/v$
filtering heuristics employed in \wcd\/ \cite{Hazelhurst08a}.
These heuristics allow us to quickly dismiss many of the edges as too
large without the need to apply the full $d^2$ algorithm (see
Sections~A.3 of the Supplementary Materials for more details).

\section{Results}

\peace\/ has been tested on both simulated and real data from NGS and
Sanger Sequencing technologies, comparing results against those
produced by the \wcd\/ clustering tool \cite{Hazelhurst08a} and the
\capthree\/ assembly tool \cite{Huang99} (the latter of which
implicitly calculates a clustering in the process of assembly).
\mc{For our simulation tests we used the {\bf ESTSim} tool to generate
  simulated Sanger sequenced transcript fragments, and the \metasim\/
  tool to generate simulated short read sequences from 454 and Illumina
  technologies \cite{Hazelhurst03,Richter2008}.  (See Supplementary
  Materials, Section~D.1, for details on the sequence size and error
  models.)  Fragments were generated from the list of 100 zebra fish
  genes used in the \wcd\/ testing \cite{Hazelhurst08a}.  Tool
  parameters were taken to match, as closely as possible, those used
  in the \wcd\/ study (see Supplementary Materials).  }

\mc{The most important method of quality assessment is {\it
    sensitivity} (the fraction of fragment pairs from the same gene
  that were} \mc{correctly clustered together).  We also look at the
  {\it Jaccard Index} (which balances sensitivity with the number of
  false positives), {\it Type 1 error} (the fraction of genes that
  were divided between clusters), and {\it Type 2 error} (the fraction
  of clusters containing two or more genes)
  \cite{Hazelhurst08a,Wang04}.  In Figure~\ref{SeJiT1T2} we plot these
  four tests as a function of error rate, and observe the almost
  identical results between \peace\/ and \wcd\/.  In
  Figure~\ref{seq_parallel} we plot the runtime for \peace\/ and
  \wcd\/, again observing almost identical results when run
  sequentially -- but significantly faster runtime for PEACE on
  multiple processor when holding the EST/processor ratio constant
  (ranging from a 65\% improvement for two processors to a 17\%
  improvement for 12 processors).}

\mc{In the course of our simulations we also investigated the memory
  footprint -- of particular concern in the large data sets commonly
  encountered.  In short, we find that, as with \wcd, the required
  memory for an execution is linear in the size of the data set --
  requiring slightly more memory than \wcd\/ and considerably less
  than \capthree.  We present relevant results in Section D.4 and
  Figure~S3 of the Supplementary Materials.}

% Appologizes to anyone working with this: the coloring would not
% presist between columns, and the break point kept changing on me.
% Modifying the \mc macro will allow for color removal without having
% to take all of this out.
\mc{We note, as a point of interest, that we can significantly improve
  the quality of \peace\/} \mc{simulation} \mc{results} \mc{through} \mc{the} \mc{increase} \mc{of}
  \mc{the} \mc{threshold} \mc{value} \mc{(see} \mc{Supplementary} \mc{Materials,} \mc{Section~D.5) --}
  \mc{achieving} \mc{a} \mc{significant} \mc{improvement} \mc{in PEACE sensitivity without an
  adverse effect on the Jaccard Index.  However, the improvement does
  not carry over to the application of real data (Supplementary
  Materials, Section E.2), where we observe a significant increase in
  the incorrect merging of clusters.  While this might be acceptable
  to a user planning to employ an assembly tool capable of breaking
  up the clusters (but unable to re-merge a split cluster), for the
  purposes of clustering assessment it makes sense base our analysis (and
  set out default values) at the lower threshold.}

\begin{figure}
  \centerline{\includegraphics[width=3.35in]{pics.d/SeJiT1T2x40.png}}
  \caption{Comparisons of Sensitivity, Jaccard Index, Type 1 error and
    Type 1 error, based on the average over 30 simulated Sanger
    Sequence ESTs sets derived from 100 zebra fish genes (see
    Supplementary Materials, Section D, for more details).  Blue/Solid
    = \peace, Green/Dash = \wcd\/ \mc{(version 0.5.1)}, Black/Dot-Dash =
    \capthree; vertical tics = 95\% confidence intervals on estimates.
    Intervals are not presented for Type 1 error due to the effective
    lack of variance.}\label{SeJiT1T2}
\end{figure}

\begin{figure}
   \centerline{\includegraphics[width=3.35in]{pics.d/seq_parallel.png}}
   \caption{Comparisons of Runtime: On left, we compare the sequential
     runtime of \peace\/ (blue) and \wcd\/ (green) on simulate sets
     ranging in size from 10K sequences to 100K sequences.  (\capthree\/
     runtime is not reported, as the time spent on clustering cannot
     be differentiated from the time spent on assembly.)  On right,
     a comparison when run in parallel, ranging the number of
     processors from 2 to 12 while holding the EST/Processor ratio
     steady at the constant 8888.  All values represent the average
     of 30 runs; vertical tics = 95\% confidence intervals on
     estimates.  All runs were done on a 3.0 GHz Intel Xeon EM64T CPU with a 2 MB
     cache and 800 MHz front side bus, model number Xeon LV 3.0 (2005).}\label{seq_parallel}
\end{figure}

\mc{In applying the tools to real data Sanger data, we used the Human
  Benchmark Data Set used to test \easycluster\/ \cite{Picardi09}, and
  the A076941 {\it Arabidopsis thaliana} data set used to test wcd\/
  \cite{Hazelhurst08a,Hazelhurst08b} (see Table~\ref{SangerReal}).
  We notice essentially identical results for \peace\/ and wcd\/ in
  quality, both significantly better than \capthree\/ in Sensitivity and Type
  1 error rate, while slightly worse in the Jaccard Index and Type 2
  error.  In runtime we see some inconsistency, with \peace\/ showing a
  60\% runtime improvement over \wcd\/ in the first data set, but requiring 20\%
  more time than \wcd\/ in the Arabidopsis data set.  (We cannot make
  a reasonable runtime comparison to \capthree, as the runtime of that
  tool reflects both clustering and assembly.)  In Table~S1 of the
  Supplementary Materials we present runtimes of several more sets,
  observing that while \peace\/ appears to be significantly 
  faster on the smaller sets, \wcd\/ does overtake it for larger
  sets.}

\begin{table*}[tbp]
\begin{center}
\begin{tabular} {| c || c || c | c | c | c | c | c | c |}
\hline
 & 
& Sensitivity & Jaccard & \begin{tabular}{c} Type 1 \\
  error \end{tabular} & \begin{tabular}{c} Type 2 \\
  error \end{tabular} &
\begin{tabular}{c} Number \\ of  \\
    Clusters \end{tabular} & \begin{tabular}{c} Number \\ of \\
    Singletons \end{tabular}
    & \begin{tabular}{c} Single \\ processor \\ runtime (s) \end{tabular}\\
\hline
\hline EasyCluster Human & \peace & 0.998 & 0.672 & 0.153 & 0.042 & 118 & 21 & 293 \\
\cline{2-9} Benchmark & \wcd & 0.998 & 0.672 & 0.144 & 0.044 & 113 & 16 & 804 \\
\cline{2-9} (111 Genes) & \capthree & 0.657 & 0.643 & 1.000 & 0.001 & 2269 & 1827 & \mbox{NA} \\
\hline
\hline WCD A076941 & \peace & 0.932 & 0.475 & 0.351 & 0.027 & 18825 & 8951 & 1166\\
\cline{2-9} Benchmark & \wcd & 0.933 & 0.476 & 0.350 &  0.027 & 18787 & 8553 & 966 \\
\cline{2-9} (13240 genes) & \capthree & 0.826 & 0.802 & 0.486 & 0.014 & 25042 & 14916 &
\mbox{NA} \\
\hline
\end{tabular}
\end{center}
\caption{Comparisons of runs on the EasyCluster human Benchmark
  Data Set and the WCD A076941 Arabidopsis thaliana data set using the
  standard quality measurements.}\label{SangerReal}
\end{table*}

% we started with a set of
% approximately $190,000$ clean Sanger ESTs derived from the {\it
%   Chlamydomonas reinhardtii} genome \cite{Liang2008}.  To compute
% sensitivity we used the {\it gmap} tool \cite{Wu05} to map the
% fragment set to the genome, taking this as our reference clustering.
% We see slight improvements in \peace\/ over \wcd\/ in both sensitivity
% and Type 1 error (both significantly outperforming \capthree).  Using
% the \mc{Arabidopsis} data set used in Hazelhurst {\it et al.}
% \cite{Hazelhurst08a} \peace\/ again shows slight improvements in
% sensitivity and Type 1 error rates, with an 18\% speedup for \peace\/ 
% (see Supplementary Materials, Section~D, for more details).

\mc{We tested the three tools on short-read data using the \metasim\/
  tool of {\it Richter et al.} \cite{Richter2008}.  Encoded into
  \metasim\/ are sequence generation and error models corresponding to
  several technologies, including the 454 short-read technology
  (producing reads ranging in size from 200bp to 370bp with a mean of
  250bp and a standard deviation of 17bp), and the Illumina short-read
  technology (producing reads of exactly 62bp).  Experiments were run
  using two sets of \wcd\/ parameters (the default parameters, and
  setting window size to 75 and threshold to 75 -- thus matching the
  equivalent \peace\/ parameters), and using default values
  of \capthree.  Only with \peace\/ were we able to get reasonable
  results, seeing a sensitivity of 0.871 for the 454 reads and
  0.583 or the shorter Illumina reads, a Jaccard Index of 0.809 and
  0.372 respectively, and other comparably figures.  (See Table~S3 in
  the Supplementary Materials.)  In short, \peace performs quite well
  for 454 reads and provides some useful information for Illumina
  reads, while \wcd\/ and \capthree\/ are unable to produce working
  results.}

\section{Conclusions}

\mc{Here we have presented \peace, a stand alone tool for the
  high-throughput clustering of transcript fragments capable of
  dealing with sequences as short as 50 bases.  \peace\/ is
  open-source and managed through a user friendly GUI that enables
  both local and remote installation and execution in sequential or
  parallel mode.  Based on a novel minimum spanning tree based
  algorithm for the clustering of the fragments by gene association,
  \peace\/ shows significant improvement in sensitivity over the
  competing \wcd\/} \mc{tool} \cite{Hazelhurst08a} \mc{when} \mc{applied} \mc{to} \mc{NGS}
  \mc{reads,} \mc{matches} \mc{\wcd\/} \mc{when} \mc{applied} \mc{to} \mc{Sanger} \mc{sequencing}
  \mc{output,} \mc{and} \mc{shows} \mc{an} \mc{order} \mc{of} \mc{magnitude} \mc{in improvement over the
  clustering performed in the course of assembly by the \capthree\/
  tool \cite{Huang99}.}

As a clustering tool based on sequence distance, \peace\/ faces
certain inherent limitations. For example, \peace\/ cannot handle
duplicate genes; like \wcd\/, it is unable to separate clusters
corresponding to genes with a greater than 88\% similarity.
Similarly, other natural biological effects (e.g. the trans-splicing of
transcripts), effects from poorly cleaned transcript data (e.g. the
failure to remove sequencing adapters or post-transcriptional
poly(A)/(T) tails), and the presence of low-complexity repeats can
cause similar effects in these clustering tools.  These problems can be
handled through the application of the assembler, and the ability to
apply any assembler to small cluster (as opposed to the data set as a
whole) results in a significant reduction in overall assembly time.

\mc{Peace can be downloaded from
\href{http://www.peace-tools.org}{www.peace-tools.org}, where we are
committed to keep maintaining and improving the tool in the future.
Meanwhile, we are developing our own MST-based assembly tool that can
seamlessly integrate with PEACE.  The underlying modular design of
PEACE offers users many possibilities to expand and incorporate the
MST algorithm for other bioinformatics application.}

\section{Acknowledgements}

Dr. Karro was funded under a PhRMA Foundation Informatics Research
Starters Grant while conducting this research.  We would also like to
acknowledge Iddo Friedberg, David Woods, Elizabeth Bikun, Jens Mueller and David
Scoville at Miami University for their help with this project.

\vspace{3mm}
% Bib TeX

\bibliography{peace.bib}

\end{document}

% LocalWords:  arallel nalysis lustering ngine Rao Moler Ozden Zhang Liang ESTs
% LocalWords:  Karro XXXXX transcriptome pre isoforms SNPs unsequenced runtime
% LocalWords:  chlamydomondomonas ESTate PaCE xsact CLU WCD Hazelhurst et al th
% LocalWords:  CHUN nalyzer Prim's MSTs subgraph MPI SourceCode ESTSim runtimes
% LocalWords:  ests Chlamydomonas Reihartdii gmap CDNA transcriptomes cDNA www
% LocalWords:  reinhartdii conifergdb DNAc bp wcd reinhardtii PhRMA Iddo fasta
% LocalWords:  Friedberg Scoville OpenMPI initio multi Screenshots nvironment
% LocalWords:  ssembly xpression GenBank dbEST pyrosequencing ESTsim sim mRNAs
% LocalWords:  NGS polyA set's GHz Xeon Bikun Arabidopsis JVM Jaccard LV GUI's
% LocalWords:  Data Set thaliana data set EasyCluster MetaSim Illumina