### 2010-05-11 papers/NAR_web_server_2009/paper.tex

\documentclass[a4,center,fleqn]{NAR}
\usepackage[pdftex,
colorlinks=true, %makes the links show up in color, rather than in a box
citecolor=black, %color of the in-text citation numbers
urlcolor=blue %color of the url links
]{hyperref}
\usepackage{wrapfig}
\bibliographystyle{unsrt}

% Enter dates of publication
\copyrightyear{}
\pubdate{}
\pubyear{}
\jvolume{}
\jissue{}

\usepackage[usenames,dvipsnames]{color}
\definecolor{red}{rgb}{1,0,0}
\newcommand{\mc}[1]{#1}

% \articlesubtype{This is the article type (optional)}

% Some shortcuts to make life easier when writing latex
\newcommand{\peace} {{\small PEACE}}
\newcommand{\wcd} {{\small WCD}}
\newcommand{\capthree} {{\small Cap3}}
\newcommand{\easycluster} {{\small EasyCluster}}
\newcommand{\metasim} {{\small MetaSim}}
\newcommand{\tgicl} {{\small TGICL}}

\begin{document}

\title{PEACE: {\underline P}arallel {\underline E}nvironment for {\underline A}ssembly
  and {\underline C}lustering of Gene {\underline E}xpression}

\author{D.M. Rao\,$^{1}$, J.C. Moler\,$^{1}$, M. Ozden\,$^1$, Y. Zhang\,$^{1}$,
  C. Liang$^{1,2*}$\, and J.E. Karro\,$^{1,3}$\footnote{to whom
    correspondence should be addressed}}

\address{$^1$ Department of Computer Science and Software Engineering, \\
  $^2$ Department of Botany, \\
  $^3$ and Department of Microbiology, Miami University, Oxford, Ohio,
  USA}

\history{Received Feb. 13, 2010 \\ Resubmitted April 21, 2010}

\maketitle

\begin{abstract}
  \mc{We present \peace, a stand alone tool for high-throughput {\it
      ab initio} clustering of transcript fragment sequences produced
    by Next Generation or Sanger Sequencing technologies.  It is
    freely available from
    \href{http://www.peace-tools.org}{www.peace-tools.org}.  Installed
    and managed through a download-able user-friendly GUI, \peace\/
    can process large data sets of transcript fragments of length 50
    bases or greater, grouping the fragments by gene associations with
    a sensitivity compatible to leading clustering tools.  Once
    clustered, the user can employ the GUI's analysis functions,
    facilitating the easy collection of statistics and allowing them
    to single out specific clusters for more comprehensive study or
    assembly.  Using a novel {\it minimum spanning tree} based
    clustering method, \peace\/ is the equal of other leading tools in
    the literature in quality with an interface making it
    accessible to any user.  It produces results of quality virtually
    identical to those of the \wcd\/ tool when applied to Sanger
    sequences, significantly improved results over \wcd\/ and \tgicl\/ when applied
    the products of Next Generation Sequencing Technology, and
    significantly improved results over \capthree\/ in both cases.  In
    short, \peace\/ provides an intuitive graphical user interface and
    a feature-rich, parallel clustering engine that proves to be a
    valuable addition to the leading cDNA clustering tools.}

\end{abstract}

\section{Introduction}

Understanding an organism's transcriptome, the set of (spliced)
transcripts expressed by genes of the organism, is a vital step in
understanding the full functional and organizational role of the
genome in the life cycle of any eukaryote.  Studying the transcriptome
has led to gene discovery, provided information on splice variants,  
and helped shed light on the biological processes both controlling and
controlled by the genome \cite{Nagaraj07}.  However, to access those
transcripts, we must deal with the fragmented data produced by
both Next Generation and traditional Sanger sequencing technology.

In the past, access to a transcriptome sequence was primarily through
the use of Expressed Sequence Tags (ESTs), single-pass cDNA sequences
derived from transcribed mRNAs and sequenced by Sanger Sequencing
technology.  More recently, Next Generation Sequencing (NGS)
technology has begun to rapidly replace Sanger Sequencing.  For
example, ESTs now being added to the GenBank dbEST are increasingly
the product of NGS technologies such as 454 pyrosequencing, which
enables the sequencing of novel and rare transcripts at a considerably
higher \cite{Cheung2006,Emrich2007}.  From a computational perspective, this
is a mixed blessing: while NGS provides immense quantities of new
information, it also provides immensely larger data sets -- and thus a
need for fast, efficient analysis algorithms.

\enlargethispage{-65.1pt}

Given a set of transcript fragments sampled from across the genome, a
necessary first step of the set's analysis is clustering: separating
the fragments according to the transcript from which they were
derived.  Frequently performed implicitly by assembly tools,
clustering the data as a ``pre-assembly'' step has a number of
advantages.  Most significantly: performing this step will
allow the application of the assembly tool to individual clusters --
saving significant amounts of time \cite{Hazelhurst08a}.

Clustering is a computationally challenging problem; the runtime and
memory requirements to cluster on the basis of pair-wise sequence
alignments make such an approach infeasible in practice.  To deal with
this, \peace\/ combines our own version of the $d^2$ alignment-free
sequence distance function \cite{Hide94} and the concept of a {\it
  minimum spanning tree} \cite{Prim57} to quickly and accurately find
clusters of ESTs expressed from the same gene without reference to a
sequenced genome.  Compared against clustering tool in the literature
\cite{Burke99,Slater00,Huang99,Parkinson02,Kalyanaraman03,Malde03,
  Pertea2003,Ptitsyn05,Hazelhurst08a,Picardi09}, \peace\/ proves to be
produce results of quality competitive with the \wcd\/ and \tgicl\/ tools
\cite{Hazelhurst08a,Pertea2003}, and more sensitive and robust than
other tools.  From the user perspective, mo tool in the literature can
match \peace\/ for ease of instillation or use, or the post-clustering
analysis tools \peace provides.

In short, \peace\/ is a computational tool for the {\it ab initio}
clustering of transcript fragments by gene association, applicable to
both NGS and traditional Sanger Sequencing technologies.  Available
through the \href{http://www.peace-tools.org}{www.peace-tools.org}
website, the \peace\/ GUI allows the user to both easily install
(locally or remotely) and run the clustering engine, as well as
enabling transparent parallel processing and providing various tools
for result analysis.

\section{\peace: Installation and Use}

The \peace\/ GUI, available as a {\small JAR} file from the \peace\/
website (\href{http://www.peace-tools.org}{www.peace-tools.org}),
can be run on any machine supporting the standard Java Virtual
Machine (JVM).  The user can employ the GUI to install
the clustering engine and perform a clustering of a data file in {\sc
  fasta} format, view an analysis of the clusters, and produce
files containing subsets of the clusters as input to assembly tools
such as \capthree\/ \cite{Huang99}.  A typical (first) use of \peace\/
must be performed in the following manner (see
Figure~\ref{fig:workflow}):

\begin{figure}
  \centerline{\includegraphics[width=3in]{pics.d/workflow.png}}
  \caption{Overview of the procedure for clustering and analysis using
    \peace.}\label{fig:workflow}
\end{figure}

\noindent {\bf Tool Installation:} (First use only.) To install the
\peace\/ clustering engine onto a local or remote machine, the user selects
from within the GUI the appropriate menu tab (Figure~\ref{screen}(a)),
which then starts an install wizard that will prompt for the
appropriate information.  Figure~\ref{screen}(b) illustrates the
request for server information; the user has chosen to install the
\peace\/ computational tool on a remote machine and is providing
the necessary connection information.  Server information is
persistent between GUI sessions, giving the user access to \peace\/
on the target machine as needed.

\begin{figure*}
  \begin{minipage}{3in}
    \includegraphics[width=3in]{screen.d/install_page_big.png}
    \centerline{\small{(a)}}
  \end{minipage}
  \begin{minipage}{2in}
    \vspace{0.47in}
    \includegraphics[width=2in]{screen.d/add_server_big.png}
    \centerline{\small{(b)}}
  \end{minipage}
  \begin{minipage}{2in}
    \vspace{0.47in}
    \includegraphics[width=2in]{screen.d/processor_assignment_big.png}
    \centerline{\small{(c)}}
  \end{minipage}

  \begin{minipage}{3in}
    \includegraphics[width=3in]{screen.d/cluster_list_big.png}
    \centerline{\small{(d)}}
  \end{minipage}
  \begin{minipage}{1in}
    \hspace{1in}
  \end{minipage}
  \begin{minipage}{3in}
    \includegraphics[width=3in]{screen.d/classifier_big.png}
    \centerline{\small{(e)}}
  \end{minipage}

  \caption{Screenshots of the \peace\/ GUI during execution, including (a)
    GUI Welcome and server installation menu; (b) setup wizard for
    installing the computational tool on a remote server; (c) execution
    wizard for starting a selected job to be executed in parallel mode;
    (d) basic cluster output; and (e) histogram view of cluster results
    and classifier editor for setting up differential expression profiles.}\label{screen}
\end{figure*}

\noindent {\bf Job Processing:} After importing the target sequence
file into the GUI, the user starts a new job by following the wizard
menus.  Figure~\ref{screen}(c) illustrates the process of specifying
the number of processors available (if running on a machine supporting
the OpenMPI protocol -- which will be determined during job
installation).  Once executed, the GUI will manage the job thread,
alert the user when the job is completed (or when the user next runs
the GUI after completion), and copy the final results back to the
local machine if necessary.

\noindent {\bf Result Analysis:} Once the resulting clusters have been
computed, the user has several options for analysis:
\begin{itemize}
\item {\bf Export:} The user can export the contents of one or more
  clusters into a FASTA format file, obtaining a subset of the
  original target file containing the sequences corresponding to the
  selected clusters ready for processing by an assembly tool
  (e.g. \capthree\/ \cite{Huang99}).
\item {\bf View Clustering}: The user may view a list of clusters,
  expanding selected clusters to a list of all individual sequences
  (illustrated in Figure~\ref{screen}(d)).
\item {\bf Classified Summary Graph:} The user may view a distribution
  of cluster sizes.  Further, the user can set up a {\it classifier},
  associating certain patterns with specific colors.  These patterns
  ware matched against the fragment header information from the
  original FASTA file, allowing the overlay of a colored cluster size
  distributions.  For example, if the sequence names contain unique
  string patterns denoting different cDNA libraries, the classifier
  can help the user to determine and visualize the differential
  expression profiles of different libraries for a given cluster.  The
  method of setting up these classifiers, and the resulting histogram,
  is illustrated in Figure~\ref{screen}(e).
\end{itemize}
Extensive documentation for the tool has been posted on the \peace\/
website, as well as links to several tutorial videos demonstrating
\peace\/ use and capabilities.

\section{Methods}

The clustering performed by \peace\/ is based on the use of minimum
spanning trees (MSTs), known to be an effective approach for narrow
band single linkage clustering \cite{Jain99,Wan08}.  Using a graph
structure to model the fragment relationships and the $d^2$ distance
measure to assign edge weights \cite{Hide94}, we can employ Prim's
algorithm \cite{Prim57} to efficiently calculate an MST from which we
can infer a high-quality clustering solution.

The $d^2$ distance measure used to assign edge weights is an
alignment-free measurement of sequence distance that can be calculated
significantly faster than a Smith-Waterman alignment \cite{Hide94}.
$d^2$ works by comparing the frequency of words (strings of a fixed
length) appearing in a limited region of each string.  Fragments
overlapping by a sufficient length will share neighborhoods of enough
similarity to ensure a small distance even in the presence of a
moderate number of base errors.  In practice we employ our own
variation of $d^2$, the {\it two-pass $d^2$ algorithm}, which
heuristically searches for a neighborhood of maximum similarity and
then finds the $d^2$ score based on that neighborhood (see
Supplementary Materials for details).

Fragment input is modeled as a weighted, undirected graph: the
fragments are represented as nodes, with $d^2$ sequence distances
assigned to the connecting edges as weights.  Conceptually, we want to
remove each edge exceeding a threshold score from the complete graph
and define our partitions by the remaining connected components.  An
edge with a large weight connects fragments which are likely
unrelated; once such edges are removed the components define a series
of overlaps.  Those fragments that can still be connected by some path
correspond to the same gene.  However, such an approach requires the
calculation of all edge weights.  That task is infeasible both in terms
of runtime and memory usage for the data set sizes we expect to process.

\peace\/ approaches the problem by generating a minimum spanning tree of the
described graph, then removing edges exceeding our threshold.  By
using Prim's algorithm we are able to calculate edge weights
on-the-fly (reducing memory requirements) and can skip the
calculation of a majority of edge distances using the $u/v$ and $t/v$
filtering heuristics employed in \wcd\/ \cite{Hazelhurst08a}.
These heuristics allow us to quickly dismiss many of the edges as too
large without the need to apply the full $d^2$ algorithm (see
Sections~A.3 of the Supplementary Materials for more details).

\section{Results}

\peace\/ has been tested on both simulated and real data from NGS and
Sanger Sequencing technologies, comparing results against those
produced by the \wcd\/ clustering tool \cite{Hazelhurst08a} and the
\capthree\/ assembly tool \cite{Huang99} (the latter of which
implicitly calculates a clustering in the process of assembly).
\mc{For our simulation tests we used the {\bf ESTSim} tool to generate
  simulated Sanger sequenced transcript fragments, and the \metasim\/
  tool to generate simulated short read sequences from 454 and Illumina
  technologies \cite{Hazelhurst03,Richter2008}.  (See Supplementary
  Materials, Section~D.1, for details on the sequence size and error
  models.)  Fragments were generated from the list of 100 zebra fish
  genes used in the \wcd\/ testing \cite{Hazelhurst08a}.  Tool
  parameters were taken to match, as closely as possible, those used
  in the \wcd\/ study (see Supplementary Materials).  }

\mc{The most important method of quality assessment is {\it
    sensitivity} (the fraction of fragment pairs from the same gene
  that were} \mc{correctly clustered together).  We also look at the
  {\it Jaccard Index} (which balances sensitivity with the number of
  false positives), {\it Type 1 error} (the fraction of genes that
  were divided between clusters), and {\it Type 2 error} (the fraction
  of clusters containing two or more genes)
  \cite{Hazelhurst08a,Wang04}.  In Figure~\ref{SeJiT1T2} we plot these
  four tests as a function of error rate, and observe the almost
  identical results between \peace\/ and \wcd\/.  In
  Figure~\ref{seq_parallel} we plot the runtime for \peace\/ and
  \wcd\/, again observing almost identical results when run
  sequentially -- but significantly faster runtime for PEACE on
  multiple processor when holding the EST/processor ratio constant
  (ranging from a 65\% improvement for two processors to a 17\%
  improvement for 12 processors).}

\mc{In the course of our simulations we also investigated the memory
  footprint -- of particular concern in the large data sets commonly
  encountered.  We find that, as with \wcd, the memory used for
  execution is linear in the size of the data set, requiring
  slightly more memory than \wcd\/ and considerably less than
  \capthree.  (See Section D.4 and Figure~S3 of the Supplementary
  Materials.}

% Appologizes to anyone working with this: the coloring would not
% presist between columns, and the break point kept changing on me.
% Modifying the \mc macro will allow for color removal without having
% to take all of this out.
\mc{We note, as a point of interest, that we can significantly improve
  the quality of \peace\/} \mc{simulation} \mc{results} \mc{through} \mc{the} \mc{increase} \mc{of}
  \mc{the} \mc{threshold} \mc{value} \mc{(see} \mc{Supplementary} \mc{Materials,} \mc{Section~D.5) --}
  \mc{achieving} \mc{a} \mc{significant} \mc{improvement} \mc{in PEACE sensitivity without an
  adverse effect on the Jaccard Index.  However, the improvement does
  not carry over to the application of real data (Supplementary
  Materials, Section E.2), where we observe a significant increase in
  the incorrect merging of clusters.  While this might be acceptable
  to a user planning to employ an assembly tool capable of breaking
  up the clusters (but unable to re-merge a split cluster), for the
  purposes of clustering assessment it makes sense to base our analysis (and
  set out default values) at the lower threshold.}

\begin{figure}
  \centerline{\includegraphics[width=3.35in]{pics.d/SeJiT1T2x40.png}}
  \caption{Comparisons of Sensitivity, Jaccard Index, Type 1 error and
    Type 1 error, based on the average over 30 simulated Sanger
    Sequence ESTs sets derived from 100 zebra fish genes (see
    Supplementary Materials, Section D, for more details).  Blue/Solid
    = \peace, Green/Dash = \wcd\/ \mc{(version 0.5.1)}, Black/Dot-Dash =
    \capthree; vertical tics = 95\% confidence intervals on estimates.}\label{SeJiT1T2}
\end{figure}

\begin{figure}
   \centerline{\includegraphics[width=3.35in]{pics.d/seq_parallel.png}}
   \caption{Comparisons of Runtime: On left, we compare the sequential
     runtime of \peace\/ (blue) and \wcd\/ (green) on simulate sets
     ranging in size from 10K sequences to 100K sequences.  (\capthree\/
     runtime is not reported, as the time spent on clustering cannot
     be differentiated from the time spent on assembly.)  On right,
     a comparison when run in parallel, ranging the number of
     processors from 2 to 12 while holding the EST/Processor ratio
     steady at the constant 8888.  All values represent the average
     of 30 runs; vertical tics = 95\% confidence intervals on
     estimates.  All runs were done on a 3.0 GHz Intel Xeon EM64T CPU with a 2 MB
     cache and a 800 MHz front side bus, model number E5520 (2005).}\label{seq_parallel}
\end{figure}

Comparisons against \tgicl\/ \cite{Pertea2003} are more complicated.
We find that when \tgicl\/ is more sensitive at lower error rates,
\peace\/ is more robust to error, and in all cases \peace\/ has an
improved Type 2 error rates and is better able to distinguish
duplicated genes.  (See Supplementary Materials Section D.6.)

In applying the tools to real data Sanger data, we used the Human
Benchmark Data Set used to test \easycluster\/ \cite{Picardi09}, and
the A076941 {\it Arabidopsis thaliana} data set used to test wcd\/
\cite{Hazelhurst08a} (see Table~\ref{SangerReal}).  We notice mixed
results, with \peace, \wcd, and \tgicl\/ showing comparable similarity,
\peace\/ and \wcd\/ showing some superiority to \tgicl\/ in Type 1 error
rate but with a poorer showing in Type 2 error, and with all three
tools coming considerably closer to the correct number of clusters
than does \capthree.  In runtime we see some inconsistency, with
\peace\/ showing a 60\% runtime improvement over \wcd\/ in the first
data set and a time comparable to \tgicl, but requiring 20\% more time
than \wcd\/ in the Arabidopsis data set and losing to \tgicl\/ by a
large margin.  In Table~S1 of the Supplementary Materials we present
runtimes of several more sets for the distance-based tool, observing
that while \peace\/ appears to be significantly faster on the smaller
sets, \wcd\/ does overtake it for larger sets.

\begin{table*}[tbp]
\begin{center}
\begin{tabular} {| c || c || c | c | c | c | c | c | c |}
\hline
 & 
& Sensitivity & Jaccard & \begin{tabular}{c} Type 1 \\
  error \end{tabular} & \begin{tabular}{c} Type 2 \\
  error \end{tabular} &
\begin{tabular}{c} Number \\ of  \\
    Clusters \end{tabular} & \begin{tabular}{c} Number \\ of \\
    Singletons \end{tabular}
    & \begin{tabular}{c} Single \\ processor \\ runtime (s) \end{tabular}\\
\hline
\hline EasyCluster Human & \peace   & 0.998 & 0.672 & 0.153 & 0.042 & 118 & 21 & 293 \\
\cline{2-9} Benchmark & \wcd        & 0.998 & 0.672 & 0.144 & 0.044 & 113 & 16 & 804 \\
\cline{2-9} (111 Genes) & \capthree & 0.657 & 0.643 & 1.000 & 0.001 & 2269 & 1827 & \mbox{NA} \\
\cline{2-9}             & \tgicl    & 0.998 & 0.949 & 0.568 & 0.018 & 221 & 86 & 278 \\
\hline
\hline A076941 & \peace           & 0.932 & 0.475 & 0.351 & 0.027 & 18825 & 8951 & 1166\\
\cline{2-9} Benchmark & \wcd          & 0.933 & 0.476 & 0.350 & 0.027 & 18787 & 8553 & 966 \\
\cline{2-9} (13240 genes) & \capthree & 0.826 & 0.802 & 0.486 & 0.014 & 25042 & 14916 & \mbox{NA} \\
\cline{2-9} & \tgicl                  & 0.939 & 0.209 & 0.401 & 0.020 & 20248 & 1065 & 425 \\
\hline
\end{tabular}
\end{center}
\caption{Comparisons of runs on the EasyCluster Human Benchmark
  Data Set and the WCD A076941 Arabidopsis thaliana data set using the
  standard quality measurements.}\label{SangerReal}
\end{table*}

% we started with a set of
% approximately $190,000$ clean Sanger ESTs derived from the {\it
%   Chlamydomonas reinhardtii} genome \cite{Liang2008}.  To compute
% sensitivity we used the {\it gmap} tool \cite{Wu05} to map the
% fragment set to the genome, taking this as our reference clustering.
% We see slight improvements in \peace\/ over \wcd\/ in both sensitivity
% and Type 1 error (both significantly outperforming \capthree).  Using
% the \mc{Arabidopsis} data set used in Hazelhurst {\it et al.}
% \cite{Hazelhurst08a} \peace\/ again shows slight improvements in
% sensitivity and Type 1 error rates, with an 18\% speedup for \peace\/ 
% (see Supplementary Materials, Section~D, for more details).

We tested the tools on short-read data using the \metasim\/ tool
of {\it Richter et al.} \cite{Richter2008}.  Encoded into \metasim\/
are sequence generation and error models corresponding to several
technologies, including the 454 short-read technology (producing reads
ranging in size from 200bp to 370bp with a mean of 250bp and a
standard deviation of 17bp), and the Illumina short-read technology
(producing reads of exactly 62bp).  In the 454 experiments \peace\/
achieved a sensitivity of 0.871, an 89.7\% improvement in sensitivity
over \wcd\/ and a 193\% improvement in sensitivity over \tgicl, with
comparable values for the Jaccard Index.  \wcd\/ was unable to handle
the Illumina data, while \peace\/ sensitivity, at a value of 0.384,
was nearly twice as good as TGICL.  (For more details, see Table~S3
in the Supplementary Materials.)  In short, \peace\/ performs quite
well for 454 reads and provides some useful information for Illumina
reads, with considerably better results than \wcd\/ or \tgicl\/.

\section{Conclusions}

Here we have presented \peace, a stand alone tool for the
high-throughput clustering of transcript fragments capable of dealing
with sequences as short as 50 bases.  \peace\/ is open-source and
managed through a user friendly GUI that enables both local and remote
installation and execution in parallel mode.  Using a novel minimum
spanning tree based algorithm for the clustering of fragments by gene
association, \peace\/ shows significant improvement in sensitivity
over the competing \wcd\/ tool and \tgicl tool when applied to NGS
reads \cite{Hazelhurst08a,Pertea2003}, matches \wcd\/ when applied to
Sanger sequencing output, and shows an order of magnitude in
improvement over the clustering performed in the course of assembly by
the \capthree\/ tool \cite{Huang99}.

As a clustering tool based on sequence distance, \peace\/ faces
certain inherent limitations. \peace\/ cannot handle duplicate genes;
like \wcd\/, it is unable to separate clusters corresponding to genes
with a greater than 88\% similarity.  Similarly, other natural
biological effects (e.g. the trans-splicing of transcripts), effects
from poorly cleaned transcript data (e.g. the failure to remove
sequencing adapters or post-transcriptional poly(A)/(T) tails), and
the presence of low-complexity repeats can cause similar effects in
these clustering tools.  These problems can be handled through the
application of the assembler, and the ability to apply any assembler
to small clusters results in a significant reduction in overall
assembly time.

Peace can be downloaded from
\href{http://www.peace-tools.org}{www.peace-tools.org}, where we are
committed to keep maintaining and improving the tool in the future.
Meanwhile, we are developing our own MST-based assembly tool that can
seamlessly integrate with PEACE.  The underlying modular design of
PEACE offers users many possibilities to expand and incorporate the
MST algorithm for other bioinformatics application.

\section{Acknowledgements}

Dr. Karro was funded under a PhRMA Foundation Informatics Research
Starters Grant while conducting this research.  We would also like to
acknowledge Iddo Friedberg, David Woods, Elizabeth Bikun, Jens Mueller and David
Scoville at Miami University for their help with this project.

\vspace{3mm}
% Bib TeX

\bibliography{peace.bib}

\end{document}

% LocalWords:  arallel nalysis lustering ngine Rao Moler Ozden Zhang Liang ESTs
% LocalWords:  Karro XXXXX transcriptome pre isoforms SNPs unsequenced runtime
% LocalWords:  chlamydomondomonas ESTate PaCE xsact CLU WCD Hazelhurst et al th
% LocalWords:  CHUN nalyzer Prim's MSTs subgraph MPI SourceCode ESTSim runtimes
% LocalWords:  ests Chlamydomonas Reihartdii gmap CDNA transcriptomes cDNA www
% LocalWords:  reinhartdii conifergdb DNAc bp wcd reinhardtii PhRMA Iddo fasta
% LocalWords:  Friedberg Scoville OpenMPI initio multi Screenshots nvironment
% LocalWords:  ssembly xpression GenBank dbEST pyrosequencing ESTsim sim mRNAs
% LocalWords:  NGS polyA set's GHz Xeon Bikun Arabidopsis JVM Jaccard LV GUI's
% LocalWords:  Data Set thaliana data set EasyCluster MetaSim Illumina tgicl