### 2011-01-10 peace/papers/EAST/east.tex
First version of the EAST manuscript (FINALLY)

\NeedsTeXFormat{LaTeX2e}[1995/12/01]
\documentclass[10pt]{bmc_article}

% Load packages
\usepackage{cite} % Make references as [1-4], not [1,2,3,4]
\usepackage{url}  % Formatting web addresses  
\usepackage{ifthen}  % Conditional 
\usepackage{multicol}   %Columns
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\urlstyle{rm}

%%%%%%%%%%%%%%%%%%%%
%% Karro macros   %%
%%%%%%%%%%%%%%%%%%%%
\newcommand{\peace} {{\small PEACE}}
\newcommand{\wcd} {{\small WCD}}
\newcommand{\capthree} {{\small Cap3}}
\newcommand{\easycluster} {{\small EasyCluster}}
\newcommand{\estsim}{{\small ESTSim}}
\newcommand{\metasim} {{\small MetaSim}}
\newcommand{\tgicl} {{\small TGICL}}
\newcommand{\east} {{\small EAST}}
\newcommand{\velvet}{{\small Velvet}}
\newcommand{\mira}{{\small Mira3}}
\newcommand{\pave} {{\small PAVE}}
\newcommand{\peast}{{\small PEACE+EAST}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%   
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %% 
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %% 
%%  submitted article.                         %% 
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{graphicx}
%\def\includegraphic{}
%\def\includegraphics{}

\setlength{\topmargin}{0.0cm}
\setlength{\textheight}{21.5cm}
\setlength{\oddsidemargin}{0cm} 
\setlength{\textwidth}{16.5cm}
\setlength{\columnsep}{0.6cm}

\newboolean{publ}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                              %%
%% You may change the following style settings  %%
%% Should you wish to format your article       %%
%% in a publication style for printing out and  %%
%% sharing with colleagues, but ensure that     %%
%% before submitting to BMC that the style is   %%
%% returned to the Review style setting.        %%
%%                                              %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%Review style settings
\newenvironment{bmcformat}{\begin{raggedright}\baselineskip20pt\sloppy\setboolean{publ}{false}}{\end{raggedright}\baselineskip20pt\sloppy}

%Publication style settings
%\newenvironment{bmcformat}{\fussy\setboolean{publ}{true}}{\fussy}

% Begin ...
\begin{document}
\begin{bmcformat}

\title{EAST: {\underline E}xpression Fragment {\underline A}ssembly from {\underline S}panning {\underline T}rees}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Ensure \and is entered between all but   %%
%% the last two authors. This will be       %%
%% replaced by a comma in the final article %%
%%                                          %%
%% Ensure there are no trailing spaces at   %% 
%% the ends of the lines                    %%     	
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author{Yuan Zhang$^1$
  \email{Yuan Zhang - zhangy9@muohio.edu}
  \and
  Dhananjai M Rao$^1$
  \email{Dhananjai Rao - raodm@muohio.edu}
  \and
  Mufit Ozden$^1$
  \email{Mufit Ozden - ozdenm@muohio.edu}
  \and
  James C Moler$^1$
  \email{James Moler - molerjc@muohio.edu}
  \and
  Chun Liang\correspondingauthor$^{1,2}$
  \email{Chun Liang\correspondingauthor - liangc@muohio.edu}
  and
  John E. Karro\correspondingauthor$^{1,3}$
  \email{John E. Karro\correspondingauthor - karroje@muohio.edu}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\address{
    \iid(1)Department of Computer Science and Software Engineering, Miami University, Oxford OHIO, USA \\
    \iid(2)Department of Botanty, Miami University, Oxford OHIO, USA \\
    \iid(3)Department of Microbiology, Miami University, Oxford OHIO, USA
}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% The Section headings here are those for  %%
%% a Research article submitted to a        %%
%% BMC-Series journal.                      %%  
%%                                          %%
%% If your article is not of this type,     %%
%% then refer to the Instructions for       %%
%% authors on http://www.biomedcentral.com  %%
%% and change the section headings          %%
%% accordingly.                             %%   
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
        % Do not use inserted blank lines (ie \\) until main body of text.
\end{abstract}

\ifthenelse{\boolean{publ}}{\begin{multicols}{2}}{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Main Body begins here                %%
%%                                          %%
%% The Section headings here are those for  %%
%% a Research article submitted to a        %%
%% BMC-Series journal.                      %%  
%%                                          %%
%% If your article is not of this type,     %%
%% then refer to the instructions for       %%
%% authors on:                              %%
%% http://www.biomedcentral.com/info/authors%%
%% and change the section headings          %%
%% accordingly.                             %% 
%%                                          %%
%% See the Results and Discussion section   %%
%% for details on how to create sub-sections%%
%%                                          %%
%% use \cite{...} to cite references        %%
%%  \cite{koon} and                         %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Introduction}
The {\it de novo} assembly of RNA transcripts is an important and
computationally challenging problem, made more difficult by the advent
of Next Generation Sequencing technologies such as 454 or Illumina
\cite{Nagaraj07,Rao10}.  Given the completion of a large-scale sequencing run of
multiple mature transcripts of a given organism, the investigator is
left with a large set of sequence fragments, but no ordering or even
transcript association information; these fragments must be clustered
by transcript then assembled before the sequencing output can be truly
useful.  If the genome has not been sequenced, this must be done {\it
  de novo} -- using only the sequence information extrapolated from
the sequences themselves.  Even when the genomic sequence is known,
there are a number of advantages to ignoring the known sequence
information and starting from scratch.  Using an existing sequence as
a template for sequencing in population studies has the potential to
introduce bias.  And when looking specifically for genes, a known
sequence may be of limited use: the target transcripts may be
undiscovered, and the splicing of introns before sequencing can
complicate the approach of mapping fragments to the genomic sequence.
In short, there is significant value in tools that can reassemble
fragment data into the original assembly without requiring access to
genomic information.

The problem of {\it de novo} transcript assembly has been addressed a
number of times in the literature, with initial approaches
concentrating on the longer reads produced by traditional Sanger
Sequencing techniques, and newer approaches looking at produces of
Next Generation Sequencing (NGS) technologies.  A list of the tools
most successfully addressing these problems include \capthree, \tgicl,
\velvet, and \mira, each excelling within their own context
\cite{Huang99,Pertea03,Chevreux04,Zerbino08}.  \capthree\/ and
\tgicl\/ use alignment-based strategies that produce high quality
results when applied to the output of Sanger sequencing technologies.
\velvet, based on the modeling of the problem with a {\it de bruijn}
graphs, was designed to handle short read sequences.  \mira\/ is a
multi-pass DNA assembler based on an overlap-layout-consensus
strategy.  [STILL NEED TO SAY SOMETHING ABOUT MIRA AND HYBRID SETS.]

In this paper we present \east, a tool for the assembly of both
Sanger, NGS and hybrid data with the capacity to incorporate quality
scores when available.  \east\/ is based on a novel minimum spanning
tree (MST) approach to compute assembly.  The tool works in tandem with
the \peace\/ clustering tool \cite{Rao10}: \peace\/ clusters
fragments by transcript and produces the initial MSTs that will serve
as the basis for the \east\/ algorithm.  In the following we present a
comparison of \east\/ against those of \capthree, \tgicl, \velvet, and
\mira on both Sanger and NGS sequences, showing that \peast\/
generates the best quality results in the fastest runtime for Sanger
sequences, and that while certain tools out-preform \peast\/ under specific
conditions, it is the only tool able to produce constant quality results
for Sanger, 454, Illumina, {\it and} hybrid data sets.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Results and Discussion %%
%%
\section*{Results and Discussion}

In assessing result quality, we must view the \peace\/ clustering tool
\cite{Rao10} and \east\/ as a combined process.  While most assembly
tools integrate these two jobs, \east\/ assumes clustering has been
preformed and requires the \peace\/ MST output.  While other cluster
tools could be substituted in to preform the actual clustering
(e.g. \wcd or \easycluster \cite{Hazelhurst08,Picardi09}), only
\peace\/ provides the requires MST output.  Hence for purposes of the
following section, we consider a single tool \peast, and assess the
quality of these results.

In order to assess the quality of the (\peace\/ +) \east\/ results, we compared
against four prominent assembly tools: \capthree, \tgicl,
\mira, and \velvet\/ \cite{Huang99,Pertea03,Chevreux04,Zerbino08}.  Tools were tested on data from three
sequencing technologies (Sanger, 454 and Illumina), using both
simulated output obtained from the \estsim\/ and \metasim\/ tools
\cite{Hazelhurst03,Richter08} and benchmark sets obtained from GenBank \cite{Benson10}.

\subsection*{Measuring Quality}
Our primary measure of quality was the {\bf normalized a-score}, based
on the {\it assembly score} (a-score) used in {\it Liang et al.}
\cite{Liang00}.  For an assembly of some original sequence, the
a-score is defined as:
$$\mbox{($2\times$sequence length) - ($5\times$number of
mismatches) - ($15 \times$number of indels)}$$
with a ``perfect'' score equal to twice the sequence length.  Our
normalized a-score modifies this definition as:
$$\frac{\mbox{($2\times$assembly length) - ($5\times$number of mismatches)
    - ($15\times$number of indels)}}{\mbox{($2\times$sequence
    length)}}$$ with a perfect score equal to 1.  This normalization
allows for the comparison of assembly quality between sequences of
different lengths. Further, using this score we can more accurately
quantify the quality of a partial reconstruction: if we are able to
reconstruct only a portion of a fragment, the normalized a-score will
take full consideration of the quality of the reconstruction, but then
bound the score by the relative size of the reconstruction.
(e.g. Perfectly reconstruction one fourth of the original sequence
will lead to an a score of $0.25$.)  In addition to a-score we also
look at: the number of contigs produced by a tool (which, in the
perfect solution, will be exactly equal to the number of transcripts
represented by the data set), normalized by the number of transcripts;
the number of singleton ESTs in the solution (i.e. ESTs unassigned to
any contig and thus, most cases, effectively thrown away), and the runtime of the tool.

\subsection*{Simulation Results}
Using simulated fragment sets allows us to both compare assemblies to
a known correct solution and control parameters in their model
(e.g. base-call error rate) to test the effect of variation.  All sets
were generated from a set of 100 zebra fish transcripts obtained from
{\it Hazelhurst et al.}  \cite{Hazelhurst08}.  For simulated Sanger
sequences we used the \estsim\/ tool, using its default model of EST
generation when not otherwise specified.  For simulated
454 and Illumina sets we used the \metasim\/ tool \cite{Richter08}, again using their
default model to simulate the errors associated with those
technologies.

\noindent {\bf Sanger Sequencing:} We first compare the assembly
quality of the tools on Sanger Sequencing, using \east, \capthree, and
\tgicl\/ (the tools designed to work with Sanger sequences), looking
at the a-score as a function of error rate for a range of coverage
depths.  For each parameter combination we applied each tool to 30
independently randomly generated simulated sets, derived from 15 zebra
fish genes, reporting the average normalized a-score.  In
Figure~\ref{sangerAscore} we look at normalized a-score as a function
of error rate, fixing coverage at 50, 30 and 10 and varying base-read
error rate from 0\% to 6\% for each of the three tools designed for
Sanger Sequencing.  While all three tools preform well when applied to
error free (i.e. impossible) sequencing technology, we find \east\/ to
be considerably more robust to base-call errors than the other tools.
Taking the mid-range coverage value of 30 as an example, at a 1\%
error rate we see \east, \capthree, and \tgicl\/ all achieving an
essentially perfect normalize a-score .  At a 3\% error rate \east\/
maintains its almost perfect normalized a-score, showing a 2\%
improvement over \tgicl\/ and a 21\% improvement over \capthree.  At a
4\% error rate \east\/ still maintains its near-perfect score (showing
a 54\% and 85\% improvement over \capthree\/ and \tgicl\/
respectively), and even at a 6\% error rate is still achieving an
normalized a-score of 0.95 (a 110\% improvement over the next closest
tool, \tgicl\/), meaning that at most 5\% of the bases were left out of
their contigs or were incorrectly reconstructed.

\begin{figure}[htb]
\centerline{\includegraphics[width=6in]{pics.d/ascore_sanger.pdf}}
\caption{Average normalized a-score as a function of base error rate for
  simulated Sanger sequencing.  \east=blue, \capthree=Black,
  \tgicl=green, \velvet=red, \mira=purple.}
\label{sangerAscore}
\end{figure}

Another useful metric is the number of contigs and number of singleton
fragments in the final result.  In Figure~\ref{contigsSanger} is a
plot of the average number of contigs per gene sequence (left column), and
the number of singletons (right column).  For the first we observe that all tools have
comparable results for lower error rates, but only \east\/ is able to
maintain the low score with increasing error rates -- generating an
perfect score in almost every case.  In minimizing the
number of singletons, we find that \east\/ does a better job than
\capthree\/ for error rates larger than $2\%$, while \tgicl does a
better job that \east\/ for error rates larger than $4\%$ to $5\%$.

\begin{figure}[htb]
\begin{minipage}{3in}
\includegraphics[width=3in]{pics.d/numContigs_sanger.pdf}
\end{minipage}
\begin{minipage}{3in}
\includegraphics[width=3in]{pics.d/numSingle_sanger.pdf}
\end{minipage}
\caption{Number of contigs (normalized to the number of transcripts)
  and number of singletons as a function of base call error
  rate for simulated Sanger sequencing.  \east=blue, \capthree=Black,
  \tgicl=green, \velvet=red, \mira=purple.  (Number of singles for
  \velvet\/ and \mira\/ not shown.)}
\label{contigsSanger}
\end{figure}

\noindent {\bf Next Generation Sequences:} In
Figure~\ref{nextgenAscore} we compare the normalized a-score for
simulated 454 and Illumina outputs (with sequence lengths averaging
250 and 62 bp respectively).  For these tests we varied coverage
levels but otherwise took the default Metasim model.  On 454 data we
see significant improvement in results quality in \east\/ for all
levels of coverage.  On Illumina data we see that \east\/ is
competitive with \velvet\/ at higher coverages, though is unable to
match that tool when coverage is sparse.

\begin{figure}[htb]
\centerline{\includegraphics[width=6in]{pics.d/ascore_nxtgen.pdf}}
\caption{Comparative normalized a-scores for 454 and Illumina sequence
  output, using the \metasim default model at different levels
  of coverage.  \east=blue, \capthree=Black, \tgicl=green,
  \velvet=red, \mira=purple.  }
\label{nextgenAscore}
\end{figure}

\noindent {\bf Hybrid:} In Figure~\ref{hybridAscore} we look at the
quality of results derived from applying each tool to a hybrid data
set of Sanger (40\%), 454 (40\%) and Illumina (20\%) sequences, again
showing normalized a-score as a function of base call error rate in
the Sanger portion.  Here we observe \east matching or outperforming
\capthree\/ and \mira\/ at moderate to high coverages.  At a coverage
level of 50, \east\/ maintains a normalized a-score hovering around
0.85 at all error levels (showing a 54\% improvement over \mira\/ at a
3\% base-call error rate).  \east\/ actually improves its performance
at a coverage level of 30 (hovering around a normalized a-score of
0.95, showing a 142\% improvement of Mira at the 3\% base-call error
rate), an oddity we attribute to the decreased absolute number of NGS
sequences with the smaller coverage.  At a coverage level of 10
\east's normalized a-score is around 0.37 -- doing considerably worse
that \mira\/ at low error rates (e.g. 0.94 v. 0.40 for error-free
bases), but showing a 52\% improvement over \mira\/ at the 3\% error
rate mark and in general showing much more robustness to error.
Remarkably, it is only in this situation that \capthree\/ is
competitive, with that tool achieving a normalized a-score of 0.62
(v. \east's 0.37) for the 3\% base call error rate, but loosing out to
\east at greater than a 4.5\% base call error rate -- a result that we
attribute to \east's problems with the 62 bp Illumina sequences (see below).

\begin{figure}[htb]
\centerline{\includegraphics[width=6in]{pics.d/ascore_hybrid.pdf}}
\caption{Comparative normalized a-scores a hybrid set of Sanger, 454
  and Illumina sequence output, varying the base call error rate for
  Sanger sequences and using the MetaSim default model for the 454 and
  Illumina sequences.  Each data set tested was composed of 40\%
  Sanger Sequences, 40\% 454 Sequences, and 20\% Illumina sequences.
  \east=blue, \capthree=Black, \tgicl=green,
  \velvet=red,\mira=purple.}
\label{hybridAscore}
\end{figure}

\noindent {\bf Tool Runtime:} As \east\/ assembly requires the
\peace\/ cluster (as opposed to the competing tools that do both
jobs), for \east\/ runtime we actually include the \peace\/
runtime. In Figures~\ref{runtime.fixed} we see the average runtime for
each of the experiments reflected in Figures
Figures~\ref{sangerAscore}-\ref{hybridAscore} for a coverage of 30,
while in Figure~\ref{runtime.coverage} we look at the average runtime
for each tool when applied to simulated data sets generated from a
single zebra fish gene at varying levels of coverage -- effectively
plotting the runtime as a function of fragment set size.  For Sanger
sequences, we see marked improvement of (\peace+)\east\/ over all
other tools for almost all coverage levels: for the smallest set
(coverage = 10) \east\/ has 1.9 speedup over \capthree\/ and a 3.9
speedup over \tgicl, while at a more moderate size (coverage=50) we
see speedups of 4.7 and 8.4.  This returns back down to 3.0 and 5.2 at the
largest set tested (coverage = 100), as can be seen in the bump around
a coverage of 85 in Figure~\ref{runtime.coverage}.  We are unable to
superficially explain this bump, but have noticed the same effect in
data sets derived from other genes as well.

\east\/ is less competitive on shorter sequence (an effect of the
$d^2$ bias of our algorithm, as discussed below), reducing its runtime
improvement for the 250 bp 454 sequences, severely reducing it for the
62 bp Illumina sequences, and also affecting the hybrid data sets (also
due to the presence of the Illumina sequence).  We argue that the high
quality of our results still makes \east\/ an appealing option for 454
and hybrid data (while acknowledging that cannot complete with
\velvet\/ on pure Illumina data), and also observe that while it has
not been implemented, the basic \east\. algorithm should easily admit
paralilization -- thus increasing the practical limit on data set size.

\begin{figure}[htb]
\centerline{\includegraphics[width=6in]{pics.d/runtime_fixedsize_sanger.pdf}}
\caption{Average runtime per trial for when generating data for
  Figures~\ref{sangerAscore}-\ref{hybridAscore}  when coverage = 30).
  Note that the exceptionally high speed of \velvet\/ on Illumina sequences make the
  red (center) bars in that plot essentially indistinguishable.
  (\peace+)\east=blue, \capthree=Black, \tgicl=green,
  \velvet=red,\mira=purple.}
\label{runtime.fixed}
\end{figure}

\begin{figure}[htb]
\centerline{\includegraphics[width=6in]{pics.d/runtime_bycoverage_all.pdf}}
\caption{The average runtime when applying each tool to a simulated
  data set derived from a specific zebrafish transcript of length
  14,625 bp.  As we decrease coverage the number of fragments drops,
  hence illustrating the relationship between tool runtime and
  fragment set size.  Each data point represents the average runtime
  from the application of the tool to at least 30 randomly generated
  simulated sets coming from the transcript at the given level of coverage.
  (\peace + )\east=blue, \capthree=Black, \tgicl=green,
  \velvet=red,\mira=purple.}
\label{runtime.coverage}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%
\section*{Conclusions}

In this manuscript we have described the \east\/ algorithm and run a
comparison of the \peast combination against the best preforming tools
in the literature.  When compared against the fine benchmark tools on
normalized a-score and runtime metrics, the \peast\/ combination of
tools prove to be both the best tool for Sanger sequencing, the
best tool overall, and by far the most robust against sequencing
error.  While certain tools may do better than it in specific circumstances,
there is no tool that does better than it on multiple sequencing
technology platforms.  While its speed is slower when applied to short
sequence fragments, the quality of its results more then compensate.

As observed in the results section, when applying \east, \capthree,
and \tgicl\/ to Sanger sequenced transcripts with normal to high error
rates, \east\/ consistently out preforms all other tools in terms of
the normalized a-score metric when applied to transcript fragments
resulting from Sanger sequencing (Figures~\ref{sangerAscore},
\ref{runtime.fixed}, and \ref{runtime.coverage}).  We also find that
the number of contigs produced is almost exactly equal to the number
of transcripts represented in the set, regardless of error rate, while
the number of singleton ESTs remains low up to about a 5\% error rate
(Figures~\ref{contigsSanger}).  The high normalized a-score indicates
that \east\/ is both reconstructing larger portions of each transcript
and doing so more faithfully, while the low number of contigs means
that we are getting complete transcripts.  Finally, runtimes for
\east\/ are considerably faster than other tools for larger sets
(Figures~\ref{runtime.fixed} and \ref{runtime.coverage}, with
(\peace+)\east\/ running a minimum of two to three times as fast as
the other tools, with speedups over \capthree\/ (the faster tool)
ranging from 1.9 to 6.0, depending on the coverage level.  In short,
when used for Sanger Sequences \east\/ is a faster tool that produces
the high quality reconstructions versus comparing tools and is
significantly more robust to error than other tools.

When involving shorter sequences, the runtime of \east\/ takes a
significant hit.  As explained in our methods section, \east\/ is
dependent on the $d^2$ sequence distance function \cite{Hide94}, whose
parameters must be adjusted when applied to smaller sequences to
preserve quality, at the cost of runtime.  We describe below our {\it
  adoptive $d^2$} algorithm that allows us to target more specific
parameters on a case-by-case basis, but large numbers of short
sequences will still lead to slower runtimes.  In our tests our
simulated 454 sequences averaged 250 bp in length, while our Illumina
sequences were uniformly 62 bp, accounting for the slightly reduced
improvement of \east, as compared to other tool runtimes, in the 454 tests and
the significant reduction for the Illumina and Hybrid test sets
containing the short Illumina sequences.

Despite the slower runtime, it can still be argued that the \east\/ is
competitive, given the quality of its output, excepting when compared
to \velvet\/ on Illumina data.  For 454 sequences, no tool comes even
close to matching the \east\/ results.  \east is able to match
\velvet for high-coverage Illumina data, though loses out at lower
coverages (and is not even remotely competitive in terms of runtime).
Looking at hybrid sets, which the \mira\/ tool is specifically geared
to address, we again find \east\/ to produce significantly superior
results at moderate to high levels of coverage.  At low coverage
levels \east\/ does suffer (another effect of the short
Illumina sequences), but its robustness then leads to
better results when the Sanger sequence portion of the set is subject
to elevated error rates.

%%%%%%%%%%%%%%%%%%
\section*{Methods}

The algorithm underlying \east\/ is based on three components: minimum
spanning trees (MSTs)\cite{Prim57}, the $d^2$ sequence distance
measure \cite{Hide94}, and standard sequence alignment
\cite{Needleman70,Smith81}.  While we would ideally be using sequence
alignment to identify overlapping ends and reassembling on that basis,
the quadratic run time of alignment algorithms coupled with the large
number of sequences makes such an approach infeasible.  To avoid excessive
applications of such algorithms, we make use MSTs to provide a {\it
  guide} through the EST set, dictating which pairs we aligned.  A
modified version of the $d^2$ measure provides the edge weights for
the underlying graph, allowing us to define the MST and also providing
some speedup to the alignment algorithm.

\east\/ is technically an assembly tool, and require clustering be
done as a pre-processing step, as well as requiring that $d^2$-based
MSTs be provided for each cluster.  Conveniently, the \peace\/
clustering tool \cite{Rao10} provides exactly this information.  While
\peace\/ an \east\/ are technically two tools, \east\/ was designed
specifically for \peace\/ output.  For purposes of this paper, it
makes sense to consider them as one.  In the following we briefly
outline the \peace\/ algorithm for clustering (as is relevant to
\east\/), and then discuss the steps taken by \east\/ for assembly of
the \peace\/ clusters.

\subsection*{Clustering}

Given a set of fragments from two or more transcripts, we must first
cluster the fragments by transcript association -- leaving us with a
single cluster of fragments for each transcript.  Details of the
\peace\/ algorithm can be found in {\it Rao et al.} \cite{Rao10}, but
the general idea is as follows: we model the problem as a weighted,
undirected graph, representing each fragment with a node and assigning
edge weights based on a $d^2$ comparison of the incident fragments
(explained in more detail below).  By employing the $u/v$ and $t/v$
filtering heuristics of {\it Hazelhurst et al.} \cite{Hazelhurst08}, we
can quickly dismiss most of the edges as irrelevant (i.e. connecting
nodes that belong in different clusters), then use Prim's algorithm to
compute an MST \cite{Prim57}.  We then remove all
edges exceeding a specified threshold value, and take each component
of the resulting forest as a cluster.  As a result \peace\/ can pass
to \east\/ a set of clusters, an MST for each cluster, and the
orientation of each fragment relative to the cluster.  Hence \east\/
has not only the clusters, but the \peace\/ MST and may quickly adjust
the sequences to ensure all are on the same strand.

\subsection*{Overlap Detection}

Given the \peace\/ generated MST for a cluster, the challenge is to
identify and align the ends of overlapping fragments without taking
the significant run time hit needed to align every pair.  We
accomplish this using two methods: we reduce the number of times an
alignment is calculated aligning only those sequences adjacent in the
MST, and we reduce the amount of  time spent per alignment by aligning
only those portions indicated by an application of a modified form of
the $d^2$ sequence distance measure.

The presence of an edge in the tree is a heuristic indication that the
incident sequences overlap; were there not sufficient sequence
similarity to indicate an overlap, the edge would have been assigned a
high $d^2$ value and would have been removed by \peace\/.  To
determine how the nodes overlap, we calculate a modified $d^2$ score
as follows.  The standard calculation of $d^2$, as defined by {\it Hide
  et al.} \cite{Hide94}, requires looking at every pair of length $w$
sub-strings (windows) from the two sequences, calculating a distance
between these two windows based on the number of shared words
(6-mers), and taking the score of the lowest-scoring window pairs as
the $d^2$ distance.  However, when looking for overlap of fragments
$s_1$ and $s_2$, we are looking only for end-overlap.  That is, we
wish to determine of the left end of $s_1$ overlaps the right end of
$s_2$, or the reverse (recalling that we know them to be on the same
strand).  Hence we need only consider the left-most and right-most
windows of $s_1$ against each window of $s_2$, reducing the number of
window comparisons from quadratic to linear in the size of $s_2$.

In Figure~\ref{fig:overlap} we illustrate the use of modified $d^2$,
using it to take two sequences (a), find the window on $s_2$ most
resembling the right-most window of $s1$ (red section of (b)),
extending to the end of $s_2$ (green section of (c)) and preforming a
{\it global} alignment on the overlapping portion.  We then compute the
{\it overlap distance} ($1 - s/l$, where $s$ is the alignment score and
$l$ is the length of the overlap segment), a value that is inversely
proportional to the probability that this is a legitimate overlap.

\begin{figure}
\centerline{\includegraphics[width=3in]{overlap_alignment.pdf}}
\caption{(a) Two sequences potentially overlapping at the ends.  (b)
  The window of $s_2$ pairing with the right-most window of $s_1$ to
  minimize the $d^2$ score.  (c) An extension of the $s_2$ window to
  the end of the sequence, and the corresponding extension on $s_1$.
  (d) A global alignment of the overlapping areas determined from
  (c).}\label{fig:overlap}
\end{figure}

We note that both \peace\/ and \east\/ make use of an {\it adaptive
  $d^2$} strategy to handle hybrid data sets.  The $d^2$ measure (and
our modification) is flexible enough to handle a range of sequence
sizes, requiring only a modification of the parameters (i.e. window
size and threshold values).  However, it runs into problems when
comparing sequences of significantly different sizes.  Comparing a 62
base Illumina read against a 1000 base Sanger read requires the use of
parameters appropriate to the shorter fragment -- which are
considerably less accurate when applied to the longer read.  Our
solution is to partition fragments by sizes (e.g. into groups of
small, medium and large fragments), and compute distances only
between segments in the name or neighboring groups.  By skipping the
comparisons between sequences of significantly different sizes we
avoid the pitfalls of such comparisons, while still picking up the
connections through transitivity.

\subsection*{Ordering and Reconstruction}

Having defined overlap distance, ordering the fragments is
straight-forward.  We pick an arbitrary node on the cluster's MST and
transverse the tree, using the overlap distance measure to check each
node against its two neighbors.  Those nodes with no identified
left-overlap are checked against nodes further out in the tree, and
any remaining with no identified left-overlap are designated as a contig
left end.

Having identified the distance of each node from its two closest
overlapping sequences, we now construct a {\it directed} graph, with
each edge pointing from a sequence to its right neighbor and weighted
with the overlap distance.  Using this graph we calculate a new
minimum spanning tree, the starting from the root of the tree
(necessarily a left-end of the contig), we transverse the tree and
assign the first base of each sequence a position relative to its
parent.  Such an assignment gives us an implicit ordering of the
fragments, by scanning through them in order and aligning each to
the next with the Smith-Waterman alignment we end up with a multiple
alignment of the fragments that allows us to derive a consensus
sequence.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Authors contributions}
    Text for this section \ldots

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
  \ifthenelse{\boolean{publ}}{\small}{}
  Text for this section \ldots

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%              
%%  Bmc_article.bst  will be used to                       %%
%%  create a .BBL file for submission, which includes      %%
%%  XML structured for BMC.                                %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %% 
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %% 
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

{\ifthenelse{\boolean{publ}}{\footnotesize}{\small}
 \bibliographystyle{bmc_article}  % Style BST file
  \bibliography{all} }     % Bibliography file (usually '*.bib' )

%%%%%%%%%%%

\ifthenelse{\boolean{publ}}{\end{multicols}}{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Figures                       %%
%%                               %%
%% NB: this is for captions and  %%
%% Titles. All graphics must be  %%
%% submitted separately and NOT  %%
%% included in the Tex document  %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Do not use \listoffigures as most will included as separate files

% \section*{Figures}
%   \subsection*{Figure 1 - Sample figure title}
%       A short description of the figure content
%       should go here.

%   \subsection*{Figure 2 - Sample figure title}
%       Figure legend text.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Tables                        %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Use of \listoftables is discouraged.
%%
% \section*{Tables}
%   \subsection*{Table 1 - Sample table title}
%     Here is an example of a \emph{small} table in \LaTeX\ using  
%     \verb|\tabular{...}|. This is where the description of the table 
%     should go. \par \mbox{}
%     \par
%     \mbox{
%       \begin{tabular}{|c|c|c|}
%         \hline \multicolumn{3}{|c|}{My Table}\\ \hline
%         A1 & B2  & C3 \\ \hline
%         A2 & ... & .. \\ \hline
%         A3 & ..  & .  \\ \hline
%       \end{tabular}
%       }
%   \subsection*{Table 2 - Sample table title}
%     Large tables are attached as separate files but should
%     still be described here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Additional Files              %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \section*{Additional Files}
%   \subsection*{Additional file 1 --- Sample additional file title}
%     Additional file descriptions text (including details of how to
%     view the file, if it is in a non-standard format or the file extension).  This might
%     refer to a multi-page table or a figure.

%   \subsection*{Additional file 2 --- Sample additional file title}
%     Additional file descriptions text.

\end{bmcformat}
\end{document}

% LocalWords:  Ozden Chun Liang Karro Botanty publ MSTs et al Hazelhurst Prim's
% LocalWords:  mers Illumina contig bmc multi LaTeX xpression ssembly rees Rao
% LocalWords:  Zhang Moler Dhananjai Mufit de novo introns NGS bruijn ABIITY bp
% LocalWords:  GenBank indels contigs TGICL Metasim MetaSim Runtime runtime pre
% LocalWords:  zebrafish runtimes ESTs paralilization