### 2025-06-29 papers/hicomb19/initial/acknowledgment.tex

\section*{ACKNOWLEDGMENT}

We thank Dr. Jens Muller for his help to run experiments on the Red
Hawk compute cluster.  We thank Dr. Scott Hazelhurst for his help with
using \q{wcd} for our experiments.

\section*{AUTHOR CONTRIBUTIONS}

D. M. Rao: A first coauthor, implemented AST and PNH in \peace,
involved in running the experiments, analyzing the data, and writing
the paper.

S. Sreeskandarajan: A first coauthor, designed PNH, involved in
running experiments, analyzing the data, and writing the paper.

C. Liang: Helped with implementation, testing, and analysis of MST
approach in \peace.