### 2025-06-29 papers/hicomb19/initial/virus_dataset.tex

\begin{table}
  \caption{Summary of Virus dataset used in experiments}
  \begin{center}
    \begin{tabular}{|l|r|c|r|p{1in}|}
      \hline

      {\bf Name} & {\bf \#Reads} & {\bf Avg Length} & {\bf Size}  & {\bf Description}  \\
           &         & {\bf (mean$\pm$ SD)} & {\bf (MB)}  &   \\ \hline

      HA   & 65,069 & 1698.8$\pm$7 & 117 & Influenza Haemmaglutinin  (HA) \\ \hline

      Dengue  & 5,286 & 10,588$\pm$181 & 55 & Degune strains (4 serotypes) \\

      \hline
    \end{tabular}\label{tab:virus-ds}
  \end{center}
\end{table}